### 2019-01-09 paper.tex
miguel initial commit of the paper structure

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{  %Calibrations of Lidars on PTU devices\\
  %Calibration of Lidars on Kinematic chain(or Robotic Arm)\\
  %calibration marker free \dots blablabla
  %Plane based calibration for Lidars on Kinematic chain (or Robotic Arm)\\
  %blablbablab \dots for 3D reconstruction\\
  %Calibration of Lidar on Robotic Arm for 3D Reconstruction\\
  %Lidar Calibration for 3D Reconstruction\\
  %Calibration method for Lidars mounted on Robotic Arms\\
  %Using Planar based features for \dots\\
%A Lidar-hand calibration method using planar based features\\
  %Calibration of a Lidar based eye-in-hand system.\\
  %A planar based optmization for lidar based eye-in-hand calibration problem\\
  Lidar to end-effector calibration approach\\ using planar based features
  *\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address}
}

\maketitle

\begin{abstract}
\dots
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

test citation \cite{Lopes2007}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}\label{sec:related_work}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed Approach}\label{sec:proposed_approach}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Future Work}\label{sec:conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\small
\bibliographystyle{IEEEtran}
\bibliography{references/refs}
}

%\begin{thebibliography}{00}
%\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
%\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
%\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
%\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
%\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
%\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
%\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
%\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2019-01-11 paper.tex
added methods straight from the thesis. Needs modification

@@
 
-test citation \cite{Lopes2007}.
+O que falar?
+\begin{itemize}
+    \item Para que pode servir esta calibração?
+    \item Onde é usada -> reconstrução de cenas, por exemplo; ou outros cenários onde a posição do laser é necessária de conhecer com elevada precisão.
+    \item 
+\end{itemize}
 
@@
 
+Pontos essenciais:
+\begin{itemize}
+    \item A calibração normalmente não é feita, por se conhecer a posição do laser por medições mecânicas;
+    \item para protótipos e desenvolvimento, por vezes a posição do Lidar não é 
+\end{itemize}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A alternative method, the Planar Based Calibration, was developed in this work to calibrate the laser scanner in this system.
+One of the key differences to the previous method (in \cref{section:radlocc-method}) is that the calibration is done only using the laser range data, and the calibration from the PTU to the laser scanner is directly found (the transformation marked in blue in \cref{figure:geometric-transformation-graph}), so no camera is required.
+This method supposes that, in a good calibration, the deviation of a point set is minimal.
+In other words, in a point set representing a planar surface, the deviation from the points to the planar surface should be the lowest, if the extrinsic calibration is correct.
+
+This method is, therefore, an optimization problem.
+For each extrinsic calibration transformation $\mathcal{T}$, corresponds a point cloud $\mathcal{P}$, following the method shown on \cref{section:point-registration}.
+This point cloud is evaluated by a cost function, which determines quantitatively how good each generated point cloud is.
+Finally, an optimizer will find the transformation $T$ that minimizes the loss function.
+Each one of these steps is described in detail next.
+
+\subsubsection{Segmentation}
+
+This calibration method uses an acquisition as its dataset, which is a significant advantage, since no calibrations patterns and no special apparatus is required, like chessboards or other markers.
+Also, a point cloud has to be generated using a estimation of the calibration transformation.
+This point cloud does not have to be geometrically accurate but the geometry should be perceivable for the plane segmentation, which is done manually prior to the calibration.
+In this work, the software CloudCompare was used to segment the point cloud into multiple planes, and the data was saved as a scalar index in each point.
+An example of a segmentation can be seen in \cref{figure:cluster-segmentation-1}, where each cluster is represented with a different color.
+
+The segmentation was done manually because most segmentation algorithms, for example the RANSAC algorithm, were not capable of achieving a reliable segmentation for the initial estimate, because the point cloud had significant deformation.
+In addition, manual segmentation is easy to do and accurate, considering that it is a one-time process.
+
+During the optimization, this segmentation serves as a blueprint for all the segmentations.
+Each point cloud is generated in the same way, so the sequence of points is always the same.
+Therefore, it is always possible to match any point on the generated point cloud to the point in the segmented point cloud, and get the corresponding cluster index for all the points.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=.8\textwidth]{cluster-segmentation-1}
+    \caption{Example of a plane segmentation, where each color represents a cluster.}
+    \label{figure:cluster-segmentation-1}
+\end{figure}
+
+\subsubsection{Cost Function}
+\label{section:calibration-cost-function}
+
+The cost function is a measure used in optimization that compares the result of a model with its expected result, and returns a value that describes the dissimilarity between the two.
+More concretely, in this calibration the cost function has two steps: the cost is computed for each cluster and then the cost of all the clusters is combined into a single value, which is the final cost of the point cloud.
+
+In an inicial step, the plane equation for each cluster is computed, using the Principal Component Analysis method, or PCA.
+First the centroid $\bar{p}$ of each plane is found, which is the same as the mean value of all the points $p: (x, y, z) \in \textbf{R}^3$:
+
+\begin{equation}
+    \bar{p} = \sum_{i}{p_i}.
+        \label{eqn:centroid-plane}
+\end{equation}
+
+Then, the covariance matrix $\mathcal{C}$ is calculated:
+
+\begin{equation}
+    \mathcal{C} = \sum_{i}{(p_i - \bar{p}) \otimes (p_i - \bar{p})} \footnote{$\otimes$ is the outer tensor product.}.
+        \label{eqn:covariance-matrix}
+\end{equation}
+
+Then, the principal axes of the plane is find by an eigen decomposition of the covariance matrix.
+The smallest eigenvalue $\lambda_3$ will be the variance $\sigma^2$ of the cluster.
+In other words, $\sigma^2$ is the mean square of the orthogonal distance of all points in the cluster to the plane.
+So, $\sigma^2$ can be a quantitative factor to measure the cost or each cluster.
+Formally, let us admit that the $\sigma^2$ has two components: the statistical error of the laser sensor $\sigma^2_{sensor}$, which is not affected by the calibration and a second component $\sigma^2_{calib}$, which depends of the calibration error.
+Thus, the idea is that, by minimizing $\sigma^2$, a exact calibration can be obtained.
+For this calibration, however, the value $\sigma$ was used instead of $\sigma^2$, which is known as the Root Mean Square Deviation, or RMS.
+Therefore, the loss of each cluster will be the $\sigma$ value.
+
+Next, the scores of the clusters are combined into a scalar value, which is the error of the point cloud.
+The method found was to, again, calculate the RMS of the values of the partial losses $\textrm{loss}_i$, according to \cref{eqn:rms}.
+This value is expected to be minimal when all the partial losses are minimal which, according to this hypothesis, corresponds to a correct calibration.
+
+\begin{equation}
+    \label{eqn:rms}
+    RMS = \sqrt{\sum_{i}^{N}{\textrm{loss}_i^2}}
+\end{equation}
+
+\subsubsection{Paramerization}
+
+The parameters in this calibration are six values that define a geometric transformation in space, which is, in the end, a transformation matrix $T$ (\cref{eqn:transformation-matrix}).
+This transformation can be decomposed into two components, a translation and a rotation.
+The translation can be represented as the vector $t = (t_x, t_y, t_z)$, and the rotation can be represented as a $3 \times 3$ rotation matrix $R$.
+Since a rotation matrix has only $3 \times 3 = 9$ elements but only 3 degrees of freedom, another parameterization has to be used to represent a rotation.
+Popular parameterization for rotations are euler angles, quaternions and axis/angle representation.
+
+\begin{equation}
+    \label{eqn:transformation-matrix}
+    T = \left[
+        \begin{array}{cccc}
+            r_{11} & r_{12} & r_{13} & t_x \\
+            r_{21} & r_{22} & r_{23} & t_y \\
+            r_{31} & r_{32} & r_{33} & t_z \\
+            0      & 0      & 0      & 1   
+        \end{array}
+    \right]
+\end{equation}
+
+However, not all representations are suitable for an optimization.
+In fact, in~\cite{hornegger99} the term fair parameterization was introduced: a parameterization is called fair, if it does not introduce more numerical sensitivity than the one inherent to the problem itself.
+Therefore, fair parameterization are a requirement for optimizations, as it increases the chances of convergence.
+For example, euler angles, which are probably the most used angle parameterization, are not suitable for optimizations~\cite{schmidt01}, because they do not yield smooth movements, each rotation is non-unique and, most notably, there are singularities, so-called \textit{gimbal-lock} singularities, where one degree of freedom is lost~\cite{schmidt01}.
+Also, quaternions are not suitable for optimizations, because quaternions have $4$ components which are constrained to an unitary length.
+Despite being a fair parameterization, quaternions introduce some complexity in the algorithm to handle this constrain, so they are not usually used for optimizations~\cite{schmidt01}.
+
+The axis/angle parameterization is the most widely used to represent a rotation in an optimization, as it is a fair parameterization and has only three components.
+Any rotation can be represented as a rotation around an axis $a$, by an angle $\theta$.
+Since $a$ only represent the direction of the rotation (hence only has 2 degrees of freedom), it can be combined with the angle $\theta$ into a single vector $\omega = \left(\omega_1, \omega_2, \omega_3\right)$, as in \cref{eqn:axis-angle}.
+
+\begin{equation}
+    \label{eqn:axis-angle}
+    \begin{aligned}
+        \theta & = |\omega| \\
+        a & = \frac{\omega}{|\omega|}
+    \end{aligned}
+\end{equation}
+
+Computing the rotation matrix from $\omega$ is done using the Rodrigues' formula~(\cref{eqn:rogriguez-1,eqn:rogriguez-2})~\cite{schmidt01}:
+
+\begin{align}
+    \label{eqn:rogriguez-1}
+    R = I + \frac{\sin \theta}{\theta} [\omega] + \frac{1 - \cos \theta}{\theta^2} [w] \\
+    \label{eqn:rogriguez-2}
+    [w] = \left[
+        \begin{array}{ccc}
+            0  & -\omega_3 & \omega_2 \\
+            \omega_3 & 0   & -\omega_1 \\
+            -\omega_2 & \omega_1 & 0 \\
+        \end{array}
+    \right],
+\end{align}
+
+\noindent
+where I is the $3\times3$ identity matrix, $\theta$ is the angle and $\omega_1$, $\omega_2$ and $\omega_3$ are the components of $\omega$.
+
+In conclusion, the parameter vector will have 6 values: 3 representing the translation ($t_1, t_2, t_3$) and 3 representing the rotation in the axis/angle representation ($r_1, r_2, r_3$).
+So, the parameter vector is shown in \cref{eqn:parameter-vector}.
+
+\begin{equation}
+    \label{eqn:parameter-vector}
+    P = \{t_1, t_2, t_3, r_1, r_2, r_3\}
+\end{equation}
+
+\subsection{First Guess}
+
+This optimization was quite robust to the initial parameters, so the first guess was always a null translation and the rotation was done doing a visual inspection of the laser scanner, using angles multiples of \SI{90}{\degree}.
+
+\subsubsection{Optimizer}
+
+The optimization is performed using the Powell's method, described in \cite{powell64}.
+This method finds a local minimum of a multi-dimensional unconstrained function, and does not require the gradient of this function (is unknown in this problem), which fits this particular optimization.
+This method is implemented in the python scientific library SciPy\footnote{See the Scipy reference in \url{https://docs.scipy.org/doc/scipy/reference/optimize.minimize-powell.html}.}.
 

### 2019-01-11 paper.tex
introduction was structured

@@
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage[utf8]{inputenc}
+
 \usepackage{cite}
@@
 \usepackage{xcolor}
+\usepackage{cleveref}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 email address}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
 }
@@
 
+%%% Notes %%%%
+
 O que falar?
@@
     \item Para que pode servir esta calibração?
-    \item Onde é usada -> reconstrução de cenas, por exemplo; ou outros cenários onde a posição do laser é necessária de conhecer com elevada precisão.
+    \item Onde é usada? reconstrução de cenas, por exemplo; ou outros cenários onde a posição do laser é necessária de conhecer com elevada precisão.
     \item 
 \end{itemize}
+
+Problem formulation
+
+What is a LRF (is 2D) -> produces 2D laser scans.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%TODO: Colocar reconstrucion para dar a motivação para o problema
+
+% 
+% Definition of the laser scanners and how the point clouds are generated.
+%
+Each laser scan is a set of range measurements taken along several directions but at the same time, all coincident with a plane, the scan plane.
+That is in fact why these equipments have  a 2D nature.
+Thus, for each scan, a 2D slice of the scene is measured.(The accumulation of these 2D slices produces the accumulated point cloud.) Since the LRF is mounted on top of an actuated kinematic chain (e.g.
+a PTU or robotic manipulator), the movement of this chain positions the LRF in different poses in the 3D space.
+Through the accumulation of the several laser scans, a dense point cloud of the scene is produced.
+
+% Explicação detalhada da acumulação e porque é impoortante a transformação do laser
+The process of accumulation, in essence, is based on the concatenation of 3D points from multiple scans.
+However, the grouping of these points requires that the points of each laser scan are transformed to a common, static, reference frame, the map frame.
+In order to achieve this, one must apply a geometric transformation to each 3D point $\bold{p}^{(i)}$ measured in scan $i$, represented in the LRF's local coordinate system $l$.
+The transformed point $\bold{q}^{(i)}$, defined in the map reference frame, is determined by:
+%
+\begin{equation}
+    \bold{q}^{(i)} = \, ^{m}\bold{T}_{e}^{(i)} \cdot ^{e}\bold{T}_{l} \cdot \bold{p}^{(i)},
+\end{equation}
+%
+\noindent where ${^{m}\bold{T}_{e}}^{(i)}$ is the transformation between the map and the end effector coordinate frames, which is dynamic and given by the process that receives a description of the kinematic chain, the joint values at the scan $i$, and computes the direct kinematics.
+In turn, $^{e}\bold{T}_{l}$ denotes the position of the LRF w.r.t the end effector.
+
+Note that, although this transformation is static, it influences the coordinates of the points $\bold{q}^{(i)}$, and therefore is crucial to the accuracy with which three dimensional objects are represented in the accumulated point cloud.
+Thus, an accurate estimation of transformation $^{e}\bold{T}_{l}$ is paramount to process of 3D reconstruction.
+We refer to this procedure of estimation the LRF to end effector transformation, as the Lidar to end-effector extrinsic calibration.
+To best of our knowledge, there is no straight forward, off-the-shelf methods for 
+conducting this calibration, as will be detailed in \cref{sec:related_work}.
+
+%TODO: Colocar aqui uma figura ilustrativa do problema
+
+In this work, we propose a novel approach to perform Lidar to end-effector extrinsic calibration.
+The approach is based on an optimization procedure which computes the transformation that generates planar point cloud sections in areas which are marked as planes.
+To this end, several existing scene structures may be used, such as walls, ceilings or pavements.
+This is an additional advantage of our proposed method: it does not require dedicated calibration targets.
+
+uses as objective function a measure of how similar to a plane portions of the point cloud are. These portions are segmente from the original point cloud using a manual one shot procedure.
+
+% By extrinsic calibration, we refer to the procedure which estimates the 6 DOF geometric transformation between the reference frame of the LRF and the reference frame of the end point of the kinematic, often referred to as the end-effector in robotic manipulators.
+% This accumulation process is the core of all 3D reconstruction systems that use LRFs.
+However this reconstruction is highly affected the estimation of the position of the LRF to the mounting point in the kinematic chain.
+% IKn this case, there are no straight forwardof the shelf methods for conducting the extrinsic calibration of the LRF.
+By extrinsic calibration, we refer to the procedure which estimates the 6dof geometric transformation between the reference frame of the LRF and the reference frame of the end point of the kinematic, often refeered to as the end-effector in robotic manipulators.
+% a plane (What is a LS?) Each LS is registered to a static frame of reference, often called the base link, as in the following equation:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Ver aqui: https://ieeexplore.ieee.org/search/searchresult.jsp?queryText=laser%20range%20finder%20calibration&highlight=true&returnFacets=ALL&returnType=SEARCH&ranges=2010_2018_Year
+
+Clustering of papers:
+
+There are some papers which use planar structures. In 
+In \cite{Kim2013}, a method for the extrinsic calibration is proposed that uses the normal verctor of two reference planes as metric for estimating the transformation. In blab
+
+% Optimizações / Grid search / analytica
+
+%In this paper, we present an algorithm for the extrinsic calibration method of rotating laser range finder (LRF).
+LRF is rotated about one axis using a motor therefore it is required to find the transformation between the motor and LRF to register scanned datum into one coordinate.
+We use calibration structure consisted of two planes to recover the transformation.
+Initial value of the transformation is set by hand and exhaustive search on the solution space is done.
+The fitness of the solution is evaluated using normal vector of the plane on the scene.
 
@@
 
+%The approach is based on an optimization procedure which computes the transformation that minimizes the distance between selected points of the accumulated point cloud and their corresponding planes.
 A alternative method, the Planar Based Calibration, was developed in this work to calibrate the laser scanner in this system.

### 2019-01-18 paper.tex
introduction

--- 2019-01-11 paper.tex
@@
 \usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\usepackage{cleveref}
+\usepackage{siunitx}
+\usepackage{hyperref}
+\usepackage[draft]{todonotes}
+\usepackage{bm}
+
+% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
 \begin{document}
 
-\title{  %Calibrations of Lidars on PTU devices\\
-  %Calibration of Lidars on Kinematic chain(or Robotic Arm)\\
-  %calibration marker free \dots blablabla
-  %Plane based calibration for Lidars on Kinematic chain (or Robotic Arm)\\
-  %blablbablab \dots for 3D reconstruction\\
-  %Calibration of Lidar on Robotic Arm for 3D Reconstruction\\
-  %Lidar Calibration for 3D Reconstruction\\
-  %Calibration method for Lidars mounted on Robotic Arms\\
-  %Using Planar based features for \dots\\
-%A Lidar-hand calibration method using planar based features\\
-  %Calibration of a Lidar based eye-in-hand system.\\
-  %A planar based optmization for lidar based eye-in-hand calibration problem\\
-  Lidar to end-effector calibration approach\\ using planar based features
-  *\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
-}
+\title{Lidar to end-effector calibration approach\\ using planar based features}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}\label{sec:introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}\label{section:introduction}
 
-O que falar?
-\begin{itemize}
-    \item Para que pode servir esta calibração?
-    \item Onde é usada -> reconstrução de cenas, por exemplo; ou outros cenários onde a posição do laser é necessária de conhecer com elevada precisão.
-    \item 
-\end{itemize}
+The Laser Range Finder, or LRF, is one of the most important sensor in the field of robotics, employed in application such as SLAM and 3D reconstruction.
+Its use is increasing due to their continuously decrease in price, new research and development and its fundamental role in emerging technologies such as autonomous driving.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}\label{sec:related_work}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% TODO: add references
+% TODO: 
+In this paper, a 2D LRF is used in a 3D reconstruction system, to produce high-detail and accurate models of real scenes.
+The common solution involves the integration of the LRF on a robotic arm, such as a PTU, to enable a full 3D reconstruction.
+However, one of the challenges of this approach is the calibration of the position of the LRF, relative to the coordinate frame of the robotic arm.
+This process is known as the LRF extrinsic calibration.
 
-Pontos essenciais:
-\begin{itemize}
-    \item A calibração normalmente não é feita, por se conhecer a posição do laser por medições mecânicas;
-    \item para protótipos e desenvolvimento, por vezes a posição do Lidar não é 
-\end{itemize}
+% TODO: add measuring instruments, explain further, add examples
+Typically, the standard procedure to do this calibration relies on a study of the geometry of such system, using a measuring instrument.
+However, this approach has several disadvantages: it is time-consuming, the instruments required are expensive and generally not easily available.
+Therefore, this solution is often not feasible for research work and prototype development.
+
+Therefore, reliable and automatic methods for the LRF extrinsic calibration are required.
+Most of these methods rely on the comparison between the structure of the reconstructed points with a known geometry, e.g.
+a marker.
+In \cite{Kim2013}, a marker formed by two perpendicular planes is used, and their points are segmented from the reconstructed point cloud.
+Then, the relation between normals of these planes are used as the objective function for the calibration.
+However, this method presents some limitations: first, ensuring that two planes are perpendicular with a small error can be very hard (carboard ); second, it is not straight-forward how to increase the number of planes; finally, by using a small marker, only a small portion of the space is occupied, which means that only a portion of the sensors' full field of view is used, which may lead to a sub-optimal result.
+
+{\color{red}
+Completar com mais estado da arte}
+
+In this paper, a new method for the extrinsic calibration of a LRF w.r.t.
+a robotic arm is presented, which makes used of the reconstructed scene geometry, more specifically, its planar regions.
+Then, the evaluation of this regions serves as the objective function for an optimization procedure, which searches the position of the LRF which minimizes the deformations of the reconstructed geometry.
+Results will show that the proposed approach is capable of generating accurate reconstructed scenes.
+
+% TODO
+The remainder of this paper will be structured as follows...
+
+\section{Problem Formulation}\label{section:problem-formulation}
+
+The LRF is a sensor that outputs several laser scans.
+Each laser scan is a set of range measurements taken along several directions but at the same time, all coincident with a plane, the scan plane.
+That is in fact why these equipments have  a 2D nature.
+Thus, for each scan, a 2D slice of the scene is measured.
+Since the LRF is mounted on top of an actuated kinematic chain (e.g.
+a PTU or robotic manipulator), the movement of this chain positions the LRF in different poses in the 3D space.
+Through the accumulation of the several laser scans, a dense point cloud of the scene is produced.
+
+The process of accumulation is, in essence, based on the concatenation of 3D points from multiple scans.
+However, the grouping of these points requires that the points of each laser scan are transformed to a common, static, reference frame, the map frame.
+In order to achieve this, one must apply a geometric transformation to each 3D point $\mathbf{p}^{(i)}$ measured in scan $i$, represented in the LRF's local coordinate system $l$.
+The transformed point $\mathbf{q}^{(i)}$, defined in the map reference frame, is determined by:
+%
+\begin{equation}
+    \mathbf{q}^{(i)} = \, ^{m}\mathbf{T}_{e}^{(i)} \cdot ^{e}\mathbf{T}_{l} \cdot \mathbf{p}^{(i)},
+\end{equation}
+%
+\noindent where ${^{m}\mathbf{T}_{e}}^{(i)}$ is the transformation between the map and the end effector coordinate frames, which is dynamic and given by the process that receives a description of the kinematic chain, the joint values at the scan $i$, and computes the direct kinematics.
+In turn, $^{e}\mathbf{T}_{l}$ denotes the position of the LRF w.r.t the end effector.
+
+Note that, although this transformation is static, it influences the coordinates of the points $\mathbf{q}^{(i)}$, and therefore is crucial to the accuracy with which three dimensional objects are represented in the accumulated point cloud.
+Thus, an accurate estimation of transformation $^{e}\mathbf{T}_{l}$ is paramount to process of 3D reconstruction.
+We refer to this procedure of estimation the LRF to end effector transformation, as the Lidar to end-effector extrinsic calibration.
+To best of our knowledge, there is no straight forward, off-the-shelf methods for 
+conducting this calibration, as will be detailed in \cref{sec:related-work}.
+
+%TODO: Colocar aqui uma figura ilustrativa do problema
+
+In this work, we propose a novel approach to perform Lidar to end-effector extrinsic calibration.
+The approach is based on an optimization procedure which computes the transformation that generates planar point cloud sections in areas which are marked as planes.
+To this end, several existing scene structures may be used, such as walls, ceilings or pavements.
+This is an additional advantage of our proposed method: it does not require dedicated calibration targets.
 
@@
 
-A alternative method, the Planar Based Calibration, was developed in this work to calibrate the laser scanner in this system.
-One of the key differences to the previous method (in \cref{section:radlocc-method}) is that the calibration is done only using the laser range data, and the calibration from the PTU to the laser scanner is directly found (the transformation marked in blue in \cref{figure:geometric-transformation-graph}), so no camera is required.
-This method supposes that, in a good calibration, the deviation of a point set is minimal.
-In other words, in a point set representing a planar surface, the deviation from the points to the planar surface should be the lowest, if the extrinsic calibration is correct.
+The working principle of this method relies on the following assunption: in a good calibration, the deviation of a measured point set w.r.t.
+to the real scene geometry should be minimal.
+This paper focuses on planar geometries because of two reasons:  they are easily segmented from the full point cloud; and it is straight-forward to compute a similarity metric between the point cloud and the expected planar geometry.
+In other words, in a point set representing a planar surface, the deviation from the points to the planar surface should be the lowest, if the calibration is correct.
 
-This method is, therefore, an optimization problem.
-For each extrinsic calibration transformation $\mathcal{T}$, corresponds a point cloud $\mathcal{P}$, following the method shown on \cref{section:point-registration}.
+This method can then be formulated as an optimization problem: for each extrinsic calibration transformation $\mathbf{T}$, corresponds a point cloud $\mathcal{P}$.
 This point cloud is evaluated by a cost function, which determines quantitatively how good each generated point cloud is.
 Finally, an optimizer will find the transformation $T$ that minimizes the loss function.
-Each one of these steps is described in detail next.
+Therefore, this method can be defined as:
+%
+\begin{equation}
+    \mathbf{T} = \underset{T}{\mathrm{argmin}} \left\{ \mathrm{\sigma}(\textbf{P}) \right\},
+\end{equation}
+%
+\noindent where % definir as variaveis
 
-\subsubsection{Segmentation}
+Falar dos vários componentes:
+
+\begin{itemize}
+    \item Parametros
+    \item Cost Fn
+    \item Optimizer
+\end{itemize}
+
+\subsection{Paramerization of the Transformation}
+
+The parameters in this calibration are six values that define a geometric transformation in space, which is, in the end, cast back to the transformation matrix $\bm{T}$.
+This transformation is decomposed into two components, a translation and a rotation.
+The translation can be represented as the translation vector $\mathbf{t} = \left[t_x, t_y, t_z\right]$, and the rotation is be represented as a $3 \times 3$ rotation matrix $\bm{R}$.
+Since a rotation matrix has only $3 \times 3 = 9$ elements but only 3 degrees of freedom, another parameterization has to be used to represent a rotation.
+Popular parameterization for rotations are euler angles, quaternions and axis/angle.
+However, not all representations are suitable for an optimization.
+In fact, in~\cite{hornegger99} the term fair parameterization was introduced: a parameterization is called fair, if it does not introduce more numerical sensitivity than the one inherent to the problem itself.
+Therefore, fair parameterizations are required for optimizations, as they increases the chances for convergence.
+For example, euler angles, which are probably the most used angle parameterizations, are not suitable~\cite{schmidt01}, because they do not yield smooth movements.
+Each rotation is non-unique and, most notably, there are singularities, so-called \textit{gimbal-lock} singularities, where one degree of freedom is lost~\cite{schmidt01}.
+Also, quaternions are not suitable for optimizations, because quaternions have $4$ components which are constrained to an unitary length.
+Despite being a fair parameterization, quaternions introduce some complexity in the algorithm to handle this constraint, so they are not usually used for optimizations~\cite{schmidt01}.
+
+The axis/angle parameterization is the most widely used to represent a rotation in optimization procedures, as it is a fair parameterization and has only three components.
+Any rotation can be represented as a rotation around an axis $a$, by an angle $\theta$.
+Since $\bm{a}$ only represents the direction of the rotation (hence only has 2 degrees of freedom), it can be combined with the angle $\theta$ into a single rotation vector $\bm{\omega} = \left[\omega_1, \omega_2, \omega_3\right]$, as follows:
+
+\begin{equation}
+    \label{eqn:axis-angle}
+    \begin{aligned}
+        \theta & = |\bm{\omega}| \\
+        \bm{a} & = \frac{\bm{\omega}}{|\bm{\omega}|}
+    \end{aligned}
+\end{equation}
+
+Computing the rotation matrix from $\omega$ is done using the Rodrigues' formula~\cite{schmidt01}:
+%
+\begin{equation}
+    \textbf{R} = \bm{I} + \frac{\sin \theta}{\theta} \bm{\Omega} + \frac{1 - \cos \theta}{\theta^2} \bm{\Omega},
+\end{equation}
+%
+\noindent
+where I is the $3\times3$ identity matrix, and $\bm{\Omega}$ is given by:
+%
+\begin{equation}
+    \bm{\Omega} = \left[
+        \begin{array}{ccc}
+            0  & -\omega_3 & \omega_2 \\
+            \omega_3 & 0   & -\omega_1 \\
+            -\omega_2 & \omega_1 & 0 \\
+        \end{array}
+    \right].
+\end{equation}
+
+In conclusion, the parameter vector to be optimized will have six values: three representing the translation $\bm{t}$ and the rotation vector $\bm{\omega}$ represented in the axis/angle representation.
+So, the parameter vector $\bm{x}$ is defined as:
+
+\begin{equation}
+    \bm{x} = \left[t_1, t_2, t_3, \omega_1, \omega_2, \omega_3\right].
+\end{equation}
+
+%TODO: Colocar dentro 
+\subsection{Segmentation}
 
@@
     \centering
-    \includegraphics[width=.8\textwidth]{cluster-segmentation-1}
+%    \includegraphics[width=.8\textwidth]{cluster-segmentation-1}
     \caption{Example of a plane segmentation, where each color represents a cluster.}
@@
 
-\subsubsection{Cost Function}
+\subsection{Cost Function}
 \label{section:calibration-cost-function}
@@
 
-\subsubsection{Paramerization}
-
-The parameters in this calibration are six values that define a geometric transformation in space, which is, in the end, a transformation matrix $T$ (\cref{eqn:transformation-matrix}).
-This transformation can be decomposed into two components, a translation and a rotation.
-The translation can be represented as the vector $t = (t_x, t_y, t_z)$, and the rotation can be represented as a $3 \times 3$ rotation matrix $R$.
-Since a rotation matrix has only $3 \times 3 = 9$ elements but only 3 degrees of freedom, another parameterization has to be used to represent a rotation.
-Popular parameterization for rotations are euler angles, quaternions and axis/angle representation.
-
-\begin{equation}
-    \label{eqn:transformation-matrix}
-    T = \left[
-        \begin{array}{cccc}
-            r_{11} & r_{12} & r_{13} & t_x \\
-            r_{21} & r_{22} & r_{23} & t_y \\
-            r_{31} & r_{32} & r_{33} & t_z \\
-            0      & 0      & 0      & 1   
-        \end{array}
-    \right]
-\end{equation}
-
-However, not all representations are suitable for an optimization.
-In fact, in~\cite{hornegger99} the term fair parameterization was introduced: a parameterization is called fair, if it does not introduce more numerical sensitivity than the one inherent to the problem itself.
-Therefore, fair parameterization are a requirement for optimizations, as it increases the chances of convergence.
-For example, euler angles, which are probably the most used angle parameterization, are not suitable for optimizations~\cite{schmidt01}, because they do not yield smooth movements, each rotation is non-unique and, most notably, there are singularities, so-called \textit{gimbal-lock} singularities, where one degree of freedom is lost~\cite{schmidt01}.
-Also, quaternions are not suitable for optimizations, because quaternions have $4$ components which are constrained to an unitary length.
-Despite being a fair parameterization, quaternions introduce some complexity in the algorithm to handle this constrain, so they are not usually used for optimizations~\cite{schmidt01}.
-
-The axis/angle parameterization is the most widely used to represent a rotation in an optimization, as it is a fair parameterization and has only three components.
-Any rotation can be represented as a rotation around an axis $a$, by an angle $\theta$.
-Since $a$ only represent the direction of the rotation (hence only has 2 degrees of freedom), it can be combined with the angle $\theta$ into a single vector $\omega = \left(\omega_1, \omega_2, \omega_3\right)$, as in \cref{eqn:axis-angle}.
-
-\begin{equation}
-    \label{eqn:axis-angle}
-    \begin{aligned}
-        \theta & = |\omega| \\
-        a & = \frac{\omega}{|\omega|}
-    \end{aligned}
-\end{equation}
-
-Computing the rotation matrix from $\omega$ is done using the Rodrigues' formula~(\cref{eqn:rogriguez-1,eqn:rogriguez-2})~\cite{schmidt01}:
-
-\begin{align}
-    \label{eqn:rogriguez-1}
-    R = I + \frac{\sin \theta}{\theta} [\omega] + \frac{1 - \cos \theta}{\theta^2} [w] \\
-    \label{eqn:rogriguez-2}
-    [w] = \left[
-        \begin{array}{ccc}
-            0  & -\omega_3 & \omega_2 \\
-            \omega_3 & 0   & -\omega_1 \\
-            -\omega_2 & \omega_1 & 0 \\
-        \end{array}
-    \right],
-\end{align}
-
-\noindent
-where I is the $3\times3$ identity matrix, $\theta$ is the angle and $\omega_1$, $\omega_2$ and $\omega_3$ are the components of $\omega$.
-
-In conclusion, the parameter vector will have 6 values: 3 representing the translation ($t_1, t_2, t_3$) and 3 representing the rotation in the axis/angle representation ($r_1, r_2, r_3$).
-So, the parameter vector is shown in \cref{eqn:parameter-vector}.
-
-\begin{equation}
-    \label{eqn:parameter-vector}
-    P = \{t_1, t_2, t_3, r_1, r_2, r_3\}
-\end{equation}
-
 \subsection{First Guess}
@@
 
-\subsubsection{Optimizer}
+\subsection{Optimizer}
 
@@
 
+% maybe describe the actual equipament used
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{itemize}
+    \item Figura com análise qualitativa (before / after)
+    \item Quantitativa (dispersão e ortogonalidade)
+\end{itemize}
 
@@
 
-{\small
+\begin{itemize}
+    \item Optimizador lento
+\end{itemize}
+
 \bibliographystyle{IEEEtran}
 \bibliography{references/refs}
-}
-
-%\begin{thebibliography}{00}
-%\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-%\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-%\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-%\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-%\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-%\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-%\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-%\end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
 

### 2019-02-05 paper.tex
finished first version

@@
 \IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+
 \usepackage[utf8]{inputenc}
@@
 \usepackage{xcolor}
-\usepackage{cleveref}
 \usepackage{siunitx}
 \usepackage{hyperref}
-\usepackage[draft]{todonotes}
+\usepackage[noabbrev, capitalize]{cleveref}
 \usepackage{bm}
-
-% \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-%    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\usepackage{todonotes}
+\usepackage{tabu}
+\usepackage{booktabs}
+\usepackage{subfig}
 
@@
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\author{\IEEEauthorblockN{Bernardo Lourenço}
+\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
+\textit{Universidade de Aveiro}\\
+Aveiro, Portugal \\
+bernardo.lourenco@ua.pt}
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\IEEEauthorblockN{Miguel Riem Oliveira}
+\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
+\textit{Universidade de Aveiro}\\
+Aveiro, Portugal \\
+mriem@ua.pt}
 \and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
+\IEEEauthorblockN{Paulo de Jesus Dias}
+\IEEEauthorblockA{\textit{IEETA} \\
+\textit{Universidade de Aveiro}\\
+Aveiro, Portugal \\
+paulo.dias@ua.pt}
 }
@@
 \begin{abstract}
-\dots
+The LRF is an essential sensor in the field of robotics, providing accurate range measurements with high angular resolution.
+This sensors are also popular in 3D reconstruction systems, mounted in kinematic chains, such as PTUs.
+However, the extrinsic transformation of the LRF has to be determined with high accuracy, otherwise the 3D reconstructions will be unusable for any subsequent procedure.
+In this work, a calibration procedure which determines this transformation is described which promises high accuracy and requires no external markers.
+This method was evaluated using real 3D reconstruction data, and proved to be successful in finding the correct transformation.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+LRF, calibration, reconstruction, point cloud, optimization
 \end{IEEEkeywords}
@@
 
-The Laser Range Finder, or LRF, is one of the most important sensor in the field of robotics, employed in application such as SLAM and 3D reconstruction.
+The Laser Range Finder, or LRF, is one of the most important sensors in the field of robotics, employed in applications such as SLAM~\cite{wang14,eyice18} and 3D reconstruction~\cite{klimentjew10,saito10}.
 Its use is increasing due to their continuously decrease in price, new research and development and its fundamental role in emerging technologies such as autonomous driving.
 
-% TODO: add references
-% TODO: 
 In this paper, a 2D LRF is used in a 3D reconstruction system, to produce high-detail and accurate models of real scenes.
@@
 
-% TODO: add measuring instruments, explain further, add examples
-Typically, the standard procedure to do this calibration relies on a study of the geometry of such system, using a measuring instrument.
+This calibration is utterly necessary because any error on it introduces a shift on the position of each point, which in turn can be seen as deformations on the point cloud resulting from the reconstruction, as can be seen in \cref{fig:deformed-pointcloud}.
+This deformations have a significant impact on subsequent point cloud processing methods, that rely on the accurate representation of the 3D environments, i.e.
+registration (for example, ICP), segmentation and feature extraction algorithms.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=8cm]{images/bad-pointcloud.png}
+    \caption{Example of deformed point cloud, product of an inaccurate calibration of the LRF. The points where this deformation is most visible is marked in red.}
+    \label{fig:deformed-pointcloud}
+\end{figure}
+
+Typically, the standard procedure to do this calibration relies on a study of the geometry of such a system, using a measuring instrument, for example, a tactile Coordinate Measuring Machine.
 However, this approach has several disadvantages: it is time-consuming, the instruments required are expensive and generally not easily available.
@@
 Therefore, reliable and automatic methods for the LRF extrinsic calibration are required.
-Most of these methods rely on the comparison between the structure of the reconstructed points with a known geometry, e.g.
+In this field, there are two big approaches, depending on the sensor which the LRF is calibrated relative to.
+The first method, and the most studied, is the calibration of the LRF relative to a camera~\cite{chen16,kassir10,zhang04}, or cameras~\cite{haeselich12}.
+In this method, the static transformation between these two sensors is determined, and relies on a correspondence between visual features captured by the camera and geometric features captured by the LRF.
+This features are usualy captured from markers placed on the field of view of both sensor, such as a chessboard, in the case of \cite{kassir10}.
+However, this methods have some disadvantages in the case of LRF mounted on robotic arms: first, they require the inclusion of a camera in the system; secondly, the calibration of the camera has to be performed; and lastly, the error of both calibrations is combined during the reconstruction, which increases the inaccuracies of the final reconstruction.
+
+Another approach is the calibration without the use of a camera, but instead the calibration is done against the robotic arm~\cite{zeng18}.
+Most of these methods rely on the comparison between the structure of the reconstructed points with known geometry, e.g.
 a marker.
-In \cite{Kim2013}, a marker formed by two perpendicular planes is used, and their points are segmented from the reconstructed point cloud.
-Then, the relation between normals of these planes are used as the objective function for the calibration.
-However, this method presents some limitations: first, ensuring that two planes are perpendicular with a small error can be very hard (carboard ); second, it is not straight-forward how to increase the number of planes; finally, by using a small marker, only a small portion of the space is occupied, which means that only a portion of the sensors' full field of view is used, which may lead to a sub-optimal result.
+In~\cite{kim13}, a marker formed by two perpendicular planes is used, and their points are segmented from the reconstructed point cloud.
+Then, the relation between the normals of these planes is used as the objective function for the calibration.
+However, this method presents some limitations: first, ensuring that two planes are perpendicular with a small error can be very hard (cardboard ); second, it is not straight-forward how to increase the number of planes; finally, by using a small marker, only a small portion of the space is occupied, which means that only a portion of the sensors' full field of view is used, which may lead to a sub-optimal result.
 
-{\color{red}
-Completar com mais estado da arte}
+In this paper, a new method for the extrinsic calibration of an LRF w.r.t.
+a robotic arm is presented, which makes use of the reconstructed scene geometry, more specifically, its planar regions.
+Then, the evaluation of this regions serves as the objective function for an optimization procedure, which searches the position of the LRF which minimizes the deformations of the reconstructed geometry.
+Results will show that the proposed approach is capable of generating accurately reconstructed scenes.
 
-In this paper, a new method for the extrinsic calibration of a LRF w.r.t.
-a robotic arm is presented, which makes used of the reconstructed scene geometry, more specifically, its planar regions.
-Then, the evaluation of this regions serves as the objective function for an optimization procedure, which searches the position of the LRF which minimizes the deformations of the reconstructed geometry.
-Results will show that the proposed approach is capable of generating accurate reconstructed scenes.
-
-% TODO
+% TODO:
 The remainder of this paper will be structured as follows...
@@
 Each laser scan is a set of range measurements taken along several directions but at the same time, all coincident with a plane, the scan plane.
-That is in fact why these equipments have  a 2D nature.
+That is in fact why these equipment have a 2D nature.
 Thus, for each scan, a 2D slice of the scene is measured.
@@
 However, the grouping of these points requires that the points of each laser scan are transformed to a common, static, reference frame, the map frame.
-In order to achieve this, one must apply a geometric transformation to each 3D point $\mathbf{p}^{(i)}$ measured in scan $i$, represented in the LRF's local coordinate system $l$.
-The transformed point $\mathbf{q}^{(i)}$, defined in the map reference frame, is determined by:
+In order to achieve this, one must apply a geometric transformation to each 3D point $\mathbf{q}^{(i)}$ measured in scan $i$, represented in the LRF's local coordinate system $l$.
+The transformed point $\mathbf{p}^{(i)}$, defined in the map reference frame, is determined by:
 %
-\begin{equation}
+\begin{equation}\label{equation:point-reconstruction}
     \mathbf{q}^{(i)} = \, ^{m}\mathbf{T}_{e}^{(i)} \cdot ^{e}\mathbf{T}_{l} \cdot \mathbf{p}^{(i)},
@@
 In turn, $^{e}\mathbf{T}_{l}$ denotes the position of the LRF w.r.t the end effector.
+In the end, the set of all the points $\mathbf{p}$ is denoted as the reconstructed point cloud $\mathcal{P}$.
 
-Note that, although this transformation is static, it influences the coordinates of the points $\mathbf{q}^{(i)}$, and therefore is crucial to the accuracy with which three dimensional objects are represented in the accumulated point cloud.
-Thus, an accurate estimation of transformation $^{e}\mathbf{T}_{l}$ is paramount to process of 3D reconstruction.
+Note that, although this transformation is static, it influences the coordinates of the points $\mathbf{p}^{(i)}$, and therefore is crucial to the accuracy with which three-dimensional objects are represented in the accumulated point cloud $\mathcal{P}$.
+Thus, an accurate estimation of transformation $^{e}\mathbf{T}_{l}$ is paramount to the process of 3D reconstruction.
 We refer to this procedure of estimation the LRF to end effector transformation, as the Lidar to end-effector extrinsic calibration.
-To best of our knowledge, there is no straight forward, off-the-shelf methods for 
-conducting this calibration, as will be detailed in \cref{sec:related-work}.
-
-%TODO: Colocar aqui uma figura ilustrativa do problema
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Proposed Approach}\label{sec:proposed_approach}
+\section{Proposed Approach}
+\label{section:proposed_approach}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The working principle of this method relies on the following assunption: in a good calibration, the deviation of a measured point set w.r.t.
+The working principle of this method relies on the following assumption: in a good calibration, the deviation of a measured point set w.r.t.
 to the real scene geometry should be minimal.
-This paper focuses on planar geometries because of two reasons:  they are easily segmented from the full point cloud; and it is straight-forward to compute a similarity metric between the point cloud and the expected planar geometry.
+This paper focuses on planar geometries because of two reasons:  they are easily segmented from the full point cloud, and it is straightforward to compute a similarity metric between the point cloud and the expected planar geometry.
 In other words, in a point set representing a planar surface, the deviation from the points to the planar surface should be the lowest, if the calibration is correct.
@@
 \begin{equation}
-    \mathbf{T} = \underset{T}{\mathrm{argmin}} \left\{ \mathrm{\sigma}(\textbf{P}) \right\},
+    \mathbf{\mathbf{x}} = \underset{\mathbf{x}}{\mathrm{argmin}} \left\{ \mathrm{f}(\mathcal{P}) \right\},
 \end{equation}
 %
-\noindent where % definir as variaveis
+\noindent where $\mathrm{f}$ is the cost function, $\mathcal{P}$ is the resulting point cloud and $x$ are the parameters that define $\mathbf{T}$.
 
-Falar dos vários componentes:
+To further explain this method, an explanation of it in each of its components will be done: the parametrization, the cost function and the optimizer.
+Each of this components are orthogonal to the others so they can be understood independently.
 
-\begin{itemize}
-    \item Parametros
-    \item Cost Fn
-    \item Optimizer
-\end{itemize}
+The optimization algorithm chosen was the Powell's method, described in \cite{powell64}.
+This method finds a local minimum of a multi-dimensional unconstrained function, and does not require the gradient of this function (is unknown in this problem), which fits this particular optimization.
+This method is implemented in the python scientific library SciPy.
+However, this method requires multiple evaluations of the objective function at each iteration, in order to calculate a approximate Jacobian matrix in the neighborhood of each point.
+This usually reflect on slow optimizations, compared to other methods, such as the Gradient Descent.
 
-\subsection{Paramerization of the Transformation}
+\subsection{Parametrization}
 
 The parameters in this calibration are six values that define a geometric transformation in space, which is, in the end, cast back to the transformation matrix $\bm{T}$.
-This transformation is decomposed into two components, a translation and a rotation.
+This transformation is decomposed into two components, a translation, and a rotation.
 The translation can be represented as the translation vector $\mathbf{t} = \left[t_x, t_y, t_z\right]$, and the rotation is be represented as a $3 \times 3$ rotation matrix $\bm{R}$.
 Since a rotation matrix has only $3 \times 3 = 9$ elements but only 3 degrees of freedom, another parameterization has to be used to represent a rotation.
-Popular parameterization for rotations are euler angles, quaternions and axis/angle.
-However, not all representations are suitable for an optimization.
-In fact, in~\cite{hornegger99} the term fair parameterization was introduced: a parameterization is called fair, if it does not introduce more numerical sensitivity than the one inherent to the problem itself.
-Therefore, fair parameterizations are required for optimizations, as they increases the chances for convergence.
-For example, euler angles, which are probably the most used angle parameterizations, are not suitable~\cite{schmidt01}, because they do not yield smooth movements.
+Popular parameterizations for rotations are Euler angles, quaternions e and axis/angle.
+However, not all representations are suitable for the optimization procedure.
+In fact, in~\cite{hornegger99} the term fair parameterization was introduced: a parameterization is called fair if it does not introduce more numerical sensitivity than the one inherent to the problem itself.
+Therefore, fair parameterizations are required for optimizations, as they increase the chances for convergence.
+For example, Euler angles, which are probably the most used angle parameterizations, are not suitable~\cite{schmidt01}, because they do not yield smooth movements.
 Each rotation is non-unique and, most notably, there are singularities, so-called \textit{gimbal-lock} singularities, where one degree of freedom is lost~\cite{schmidt01}.
-Also, quaternions are not suitable for optimizations, because quaternions have $4$ components which are constrained to an unitary length.
+Also, quaternions are not suitable for optimizations, because quaternions have $4$ components which are constrained to a unitary length.
 Despite being a fair parameterization, quaternions introduce some complexity in the algorithm to handle this constraint, so they are not usually used for optimizations~\cite{schmidt01}.
@@
 
-%TODO: Colocar dentro 
-\subsection{Segmentation}
+This optimization was quite robust to the initial parameters, so the first guess was always a null translation and the rotation was done doing a visual inspection of the laser scanner, using angles multiples of \SI{90}{\degree}.
 
-This calibration method uses an acquisition as its dataset, which is a significant advantage, since no calibrations patterns and no special apparatus is required, like chessboards or other markers.
-Also, a point cloud has to be generated using a estimation of the calibration transformation.
-This point cloud does not have to be geometrically accurate but the geometry should be perceivable for the plane segmentation, which is done manually prior to the calibration.
-In this work, the software CloudCompare was used to segment the point cloud into multiple planes, and the data was saved as a scalar index in each point.
-An example of a segmentation can be seen in \cref{figure:cluster-segmentation-1}, where each cluster is represented with a different color.
+\subsection{Cost Function}
 
-The segmentation was done manually because most segmentation algorithms, for example the RANSAC algorithm, were not capable of achieving a reliable segmentation for the initial estimate, because the point cloud had significant deformation.
+The cost function is a function used in optimization procedures that measures the difference of the result of a model and its expected result.
+It then yield a value that quantitatively describes the dissimilarity between the two.
+In this case, the cost function evaluates the distance of each point $\mathbf{p}$ in the point cloud $\mathcal{P}$ to their respective plane.
+The greater this value, the greater the deformation of the point cloud, and thus the worst the calibration is.
+This objective is, then, to find the set of parameters that minimize the cost function.
+
+As an initial step, the segmentation of the point cloud is performed.
+The segmentation is a procedure that divides the point cloud $\mathcal{P}$ into multiple clusters $\mathcal{C}_i$, each one correspondent to different planes on the scene.
+This segmentation is done prior to the calibration procedure, using the point cloud reconstructed using an initial estimate.
+Also, this segmentation was done manually because most segmentation algorithms, for example the RANSAC algorithm, were not capable of achieving a reliable segmentation, because the point cloud had significant deformation.
 In addition, manual segmentation is easy to do and accurate, considering that it is a one-time process.
 
-During the optimization, this segmentation serves as a blueprint for all the segmentations.
+During the optimization, this segmentation serves as a blueprint for all segmentations.
 Each point cloud is generated in the same way, so the sequence of points is always the same.
@@
     \centering
-%    \includegraphics[width=.8\textwidth]{cluster-segmentation-1}
+    \includegraphics[width=8cm]{images/segmented-pointcloud}
     \caption{Example of a plane segmentation, where each color represents a cluster.}
@@
 
-\subsection{Cost Function}
-\label{section:calibration-cost-function}
-
-The cost function is a measure used in optimization that compares the result of a model with its expected result, and returns a value that describes the dissimilarity between the two.
-More concretely, in this calibration the cost function has two steps: the cost is computed for each cluster and then the cost of all the clusters is combined into a single value, which is the final cost of the point cloud.
-
-In an inicial step, the plane equation for each cluster is computed, using the Principal Component Analysis method, or PCA.
+After the segmentation, the calibration procedure can be performed.
+In an initial step, the plane equation for each point cluster is computed, using the Principal Component Analysis method, or PCA.
 First the centroid $\bar{p}$ of each plane is found, which is the same as the mean value of all the points $p: (x, y, z) \in \textbf{R}^3$:
@@
 
-Then, the principal axes of the plane is find by an eigen decomposition of the covariance matrix.
+Then, the principal axes of the plane is find by an eigen-decomposition of the covariance matrix.
 The smallest eigenvalue $\lambda_3$ will be the variance $\sigma^2$ of the cluster.
@@
 
-Next, the scores of the clusters are combined into a scalar value, which is the error of the point cloud.
-The method found was to, again, calculate the RMS of the values of the partial losses $\textrm{loss}_i$, according to \cref{eqn:rms}.
-This value is expected to be minimal when all the partial losses are minimal which, according to this hypothesis, corresponds to a correct calibration.
-
+Finally, the scores of the clusters are combined into a scalar value, which is the error of the point cloud.
+The method found was to, again, calculate the RMS of the values of the partial losses of each cluster $\sigma_i$, according to:
+%
 \begin{equation}
-    \label{eqn:rms}
-    RMS = \sqrt{\sum_{i}^{N}{\textrm{loss}_i^2}}
+    \textrm{f} = \sqrt{\sum_{i}^{N}{\sigma_i^2}}.
 \end{equation}
-
-\subsection{First Guess}
-
-This optimization was quite robust to the initial parameters, so the first guess was always a null translation and the rotation was done doing a visual inspection of the laser scanner, using angles multiples of \SI{90}{\degree}.
-
-\subsection{Optimizer}
-
-The optimization is performed using the Powell's method, described in \cite{powell64}.
-This method finds a local minimum of a multi-dimensional unconstrained function, and does not require the gradient of this function (is unknown in this problem), which fits this particular optimization.
-This method is implemented in the python scientific library SciPy\footnote{See the Scipy reference in \url{https://docs.scipy.org/doc/scipy/reference/optimize.minimize-powell.html}.}.
-
-% maybe describe the actual equipament used
+%
+\noindent This value is expected to be minimal when all the partial losses are minimal which, according to this hypothesis, corresponds to a correct calibration.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}\label{sec:results}
+\section{Results}
+\label{section:results}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+In this section, the proposed method for the calibration will be evaluated using acquired data from a real scene.
+The infraestructure used to acquire the data, as well as a description of the data will be done.
+Afterward, both a qualitative and quantitative analisys of the results will be performed, to determine how much the geometry of the calibrated point cloud improved, compared to the uncalibrated point cloud.
+
+The hardware infrastructure used in this work to acquire the data was the 3D scanner \textit{lemonbot}, show in \cref{figure:lemonbot}.
+The principal components of this 3D scanner are the PTU \textit{FLIR PTU-D46} and the LRF \textit{Hokuyo UTM30LX}.
+The LRF was mounted on the PTU, to have the scan plane in the vertical direction.
+The LRF can cover \SI{270}{\degree} with an angular resolution of \SI{0.25}{\degree}.
+The PTU has an angular resolution of \SI{0.0032}{\degree}, so the error in the transformation $^{m}\mathbf{T}_{e}$ (see \cref{equation:point-reconstruction}) is negligeble.
+This 3D scanner was programmed using the \textit{ROS} framework, which is handles the hardware drivers, the transformation graph and the recording of the scanning data.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=8cm]{images/lemonbot-head}
+    \caption{The \textit{lemonbot} 3D scanner, used in this work. The LRF \textit{Hokuyo UTM30} is marked in red and the \textit{FLIR PTU-D46} is marked in green.}
+    \label{figure:lemonbot}
+\end{figure}
+
+For the verification, a set of 3 datasets was acquired from a single room, which corresponds to 3 point clouds.
+The acquisitions were performed in different positions in the room.
+The room was not cluttered and the ceiling, walls and floor are all perpendicular and known to be planar surfaces.
+This was ensured in order to minimize the segmentation and calibration errors.
+
+The optimization procedure was successful in minimizing the objective function in every dataset, as seen in \cref{table:iteration-results}.
+The final results a qualitative improvement in comparison to the uncalibrated point cloud, as show in \cref{figure:visual-comparison}.
+The deformation found in the uncalibrated point cloud are almost imperceptible in the calibrated point cloud.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=8cm]{images/calibration-evolution.png}
+    \caption{Graph of the evolution of the cost function during the optimization process. As seen, the cost function decreased roughly 50 times.}
+    \label{table:iteration-results}
+\end{figure}
+
+\begin{figure*}
+    \centering
+
+    \centering
+    \subfloat{\includegraphics[width = 8cm]{images/example-pointcloud-2-bad}} \qquad
+    \subfloat{\includegraphics[width = 8cm]{images/example-pointcloud-2-good}}
+
+    \centering
+    \subfloat{\includegraphics[width = 8cm]{images/example-pointcloud-3-bad}} \qquad
+    \subfloat{\includegraphics[width = 8cm]{images/example-pointcloud-3-good}}
+
+    \centering
+    \subfloat{\includegraphics[height = 6cm]{images/example-pointcloud-1-bad}} \qquad
+    \subfloat{\includegraphics[height = 6cm]{images/example-pointcloud-1-good}}
+
+    \caption{Side-by-side comparison of the uncalibrated (on the left) and calibrated (on the right) point clouds.
+The most noticeable distortions are marked in red, which did not appear in the calibrated point clouds.
+This indicated that the method successfully reduced the distortions in the point clouds.}
+    \label{figure:visual-comparison}
+\end{figure*}
+
+Despite the good qualitative result shown prior, a further quantitative analysis was performed in order to evaluate the method. The results were prepared in the following steps:
+
 \begin{itemize}
-    \item Figura com análise qualitativa (before / after)
-    \item Quantitativa (dispersão e ortogonalidade)
+    \item A planar surface was extracted from both the calibrated and uncalibrated point cloud.
+    \item The fitting plane was calculated for each surface, and the distance between the points to the surface was computed.
+    \item The statistic deviation of the distances was calculated.
 \end{itemize}
+
+This method was performed for 3 planes, and the segments can be seen in \cref{table:quantitative-results}.
+As seen the statistical deviation is in fact smaller in the calibrated point cloud.
+As an alternative view, in \cref{figure:deviation-histogram}, the histogram and distribution of the signed deviation is shown for both the point clouds.
+As shown, the deviations of the calibrated point cloud are less scattered than the deviations of the uncalibrated point cloud.
+This results indicate that this calibration method in fact improves the results.
+
+\begin{table}
+    \caption{Comparison between the standard deviation and mean of the distances of the points to the plane for both the calibrated and uncalibrated point clouds.}
+    \begin{tabu}{X[l] X[c] X[c]}
+        \toprule
+        Point Cloud & $\mu$ (average) & $\sigma$ (std. dev.) \\
+        \midrule
+        Calibrated     & 0.0495 & 0.2589 \\
+        Uncalibrated   & 0.0025 & 1.1211 \\
+        \bottomrule
+    \end{tabu}
+
+    \label{table:quantitative-results}
+\end{table}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=8cm]{images/pointclouds-histogram.png}
+    \caption{Distribution of the distances of the points to the plane. The uncalibrated points is a wider than the calibrated points, which indicates an improved calibration.}
+    \label{figure:deviation-histogram}
+\end{figure}
 
@@
 
-\begin{itemize}
-    \item Optimizador lento
-\end{itemize}
+This paper presents a calibration procedure to determine the extrinsic transformation of LRF mounted on a robotic arm.
+This transformation is uttermost necessary for multiple applications, most noticeable, 3D reconstruction systems.
+The incorrect calibration of this systems results in artifacts and deformations on the final reconstructions, which invalidates any successive work in this point clouds.
+The proposed method is based on the measurement of the deformations of reconstructed planar surfaces, and through an optimization procedure, this deformations are minimized.
+The method was proved to be effective, through a visual inspection of the resulting point cloud and through a statistical evaluation of the reconstructed geometries.
 

### 2019-02-08 paper.tex
revision 2

@@
 \usepackage{bm}
-\usepackage{todonotes}
 \usepackage{tabu}
@@
 \usepackage{subfig}
+\usepackage{bookmark}
 
@@
 
-\title{Lidar to end-effector calibration approach\\ using planar based features}
+\title{
+    2D-LRF to PTU calibration \\
+    using planar features
+}
 
@@
 \and
+\IEEEauthorblockN{Paulo Dias}
+\IEEEauthorblockA{\textit{IEETA} \\
+\textit{Universidade de Aveiro}\\
+Aveiro, Portugal \\
+paulo.dias@ua.pt}
+\and
 \IEEEauthorblockN{Miguel Riem Oliveira}
@@
 mriem@ua.pt}
-\and
-\IEEEauthorblockN{Paulo de Jesus Dias}
-\IEEEauthorblockA{\textit{IEETA} \\
-\textit{Universidade de Aveiro}\\
-Aveiro, Portugal \\
-paulo.dias@ua.pt}
 }
@@
 \begin{abstract}
-The LRF is an essential sensor in the field of robotics, providing accurate range measurements with high angular resolution.
+The Laser Range Finder, or LRF, is an essential sensor in the field of robotics, providing accurate range measurements with high angular resolution.
 This sensors are also popular in 3D reconstruction systems, mounted in kinematic chains, such as PTUs.
@@
 
-The Laser Range Finder, or LRF, is one of the most important sensors in the field of robotics, employed in applications such as SLAM~\cite{wang14,eyice18} and 3D reconstruction~\cite{klimentjew10,saito10}.
+The Laser Range Finder, or LRF, is one of the most important sensors in the field of robotics, employed in applications such as Simultaneous Localization and Mapping~\cite{wang14,eyice18} and 3D reconstruction~\cite{saito10}.
 Its use is increasing due to their continuously decrease in price, new research and development and its fundamental role in emerging technologies such as autonomous driving.
 
-In this paper, a 2D LRF is used in a 3D reconstruction system, to produce high-detail and accurate models of real scenes.
-The common solution involves the integration of the LRF on a robotic arm, such as a PTU, to enable a full 3D reconstruction.
-However, one of the challenges of this approach is the calibration of the position of the LRF, relative to the coordinate frame of the robotic arm.
-This process is known as the LRF extrinsic calibration.
+In this paper, a 2D-LRF is used in a 3D reconstruction system, to produce high-detail and accurate models of real scenes.
+Using a 2D-LRF is this system has the advantage of being several times cheaper than commercial 3D laser scanners~\cite{dias06}.
+The common solution involves the integration of the LRF on a turntable~\cite{maurelli09} or a Pan-tilt unit, or PTU~\cite{klimentjew09}, to enable a full 3D reconstruction.
+One of the challenges of this approach is the calibration of the position of the LRF, relative to the coordinate frame of a dynamic kinematic chain, such as a PTU.
+This process is known as the LRF-PTU extrinsic calibration.
 
-This calibration is utterly necessary because any error on it introduces a shift on the position of each point, which in turn can be seen as deformations on the point cloud resulting from the reconstruction, as can be seen in \cref{fig:deformed-pointcloud}.
+This calibration is utterly necessary because any error introduces a shift on the position of each point, which in turn can be seen as deformations on the point cloud resulting from the aquisition.
+Such deformation can be seen in \cref{fig:deformed-pointcloud}, which is shown as multiple missalign planes.
 This deformations have a significant impact on subsequent point cloud processing methods, that rely on the accurate representation of the 3D environments, i.e.
@@
 
-Typically, the standard procedure to do this calibration relies on a study of the geometry of such a system, using a measuring instrument, for example, a tactile Coordinate Measuring Machine.
+Typically, a possible solution to do this calibration relies on a study of the geometry of such a system, using a measuring instrument, for example, a tactile Coordinate Measuring Machine.
 However, this approach has several disadvantages: it is time-consuming, the instruments required are expensive and generally not easily available.
-Therefore, this solution is often not feasible for research work and prototype development.
 
 Therefore, reliable and automatic methods for the LRF extrinsic calibration are required.
-In this field, there are two big approaches, depending on the sensor which the LRF is calibrated relative to.
+In this field, there are two main approaches, depending on the sensor which the LRF is calibrated relative to.
 The first method, and the most studied, is the calibration of the LRF relative to a camera~\cite{chen16,kassir10,zhang04}, or cameras~\cite{haeselich12}.
 In this method, the static transformation between these two sensors is determined, and relies on a correspondence between visual features captured by the camera and geometric features captured by the LRF.
-This features are usualy captured from markers placed on the field of view of both sensor, such as a chessboard, in the case of \cite{kassir10}.
+This features might be markers placed on the field of view of both sensor, such as a chessboard, in the case of \cite{kassir10}.
 However, this methods have some disadvantages in the case of LRF mounted on robotic arms: first, they require the inclusion of a camera in the system; secondly, the calibration of the camera has to be performed; and lastly, the error of both calibrations is combined during the reconstruction, which increases the inaccuracies of the final reconstruction.
 
-Another approach is the calibration without the use of a camera, but instead the calibration is done against the robotic arm~\cite{zeng18}.
-Most of these methods rely on the comparison between the structure of the reconstructed points with known geometry, e.g.
-a marker.
+Another approach is to calibrate the 2D-LRF directly against a moving platform~\cite{zeng18}.
+Most of these methods rely on the comparison between the structure of the reconstructed points with a object with a known 3D geometry, such as a plane, a ball~\cite{pereira16} or a cone~\cite{almeida12}.
 In~\cite{kim13}, a marker formed by two perpendicular planes is used, and their points are segmented from the reconstructed point cloud.
 Then, the relation between the normals of these planes is used as the objective function for the calibration.
-However, this method presents some limitations: first, ensuring that two planes are perpendicular with a small error can be very hard (cardboard ); second, it is not straight-forward how to increase the number of planes; finally, by using a small marker, only a small portion of the space is occupied, which means that only a portion of the sensors' full field of view is used, which may lead to a sub-optimal result.
+However, this method presents some limitations: first, ensuring that two planes are perpendicular with a small error can be very hard (it requires a well-built marker); second, it is not straight-forward how to increase the number of planes; finally, by using a small marker, only a small portion of the space is occupied, which means that only a portion of the sensors' full field of view is used, which may lead to a sub-optimal result.
 
-In this paper, a new method for the extrinsic calibration of an LRF w.r.t.
-a robotic arm is presented, which makes use of the reconstructed scene geometry, more specifically, its planar regions.
-Then, the evaluation of this regions serves as the objective function for an optimization procedure, which searches the position of the LRF which minimizes the deformations of the reconstructed geometry.
-Results will show that the proposed approach is capable of generating accurately reconstructed scenes.
-
-% TODO:
-The remainder of this paper will be structured as follows...
+% TODO: colocar qualquer coisa aqui...
 
@@
 
-The LRF is a sensor that outputs several laser scans.
-Each laser scan is a set of range measurements taken along several directions but at the same time, all coincident with a plane, the scan plane.
+The 2D-LRF is a sensor that outputs several laser scans.
+Each laser scan is a set of range measurements taken along several directions all coincident with a plane: the scanning plane.
 That is in fact why these equipment have a 2D nature.
 Thus, for each scan, a 2D slice of the scene is measured.
-Since the LRF is mounted on top of an actuated kinematic chain (e.g.
-a PTU or robotic manipulator), the movement of this chain positions the LRF in different poses in the 3D space.
+Since the LRF is mounted on top of an actuated kinematic chain (the PTU), the movement of this chain positions the LRF in different poses in the 3D space.
 Through the accumulation of the several laser scans, a dense point cloud of the scene is produced.
@@
 However, the grouping of these points requires that the points of each laser scan are transformed to a common, static, reference frame, the map frame.
-In order to achieve this, one must apply a geometric transformation to each 3D point $\mathbf{q}^{(i)}$ measured in scan $i$, represented in the LRF's local coordinate system $l$.
+In order to achieve it, one must apply a geometric transformation to each 3D point $\mathbf{q}^{(i)}$ measured in scan $i$, represented in the LRF's local coordinate system $l$.
 The transformed point $\mathbf{p}^{(i)}$, defined in the map reference frame, is determined by:
@@
 %
-\noindent where ${^{m}\mathbf{T}_{e}}^{(i)}$ is the transformation between the map and the end effector coordinate frames, which is dynamic and given by the process that receives a description of the kinematic chain, the joint values at the scan $i$, and computes the direct kinematics.
-In turn, $^{e}\mathbf{T}_{l}$ denotes the position of the LRF w.r.t the end effector.
+\noindent where ${^{m}\mathbf{T}_{e}}^{(i)}$ is the transformation between the map and the PTU coordinate frames, which is dynamic and given by the process that receives a description of the kinematic chain, the joint values at the scan $i$, and computes the direct kinematics.
+In turn, $^{e}\mathbf{T}_{l}$ denotes the position of the LRF w.r.t the PTU.
 In the end, the set of all the points $\mathbf{p}$ is denoted as the reconstructed point cloud $\mathcal{P}$.
@@
 Thus, an accurate estimation of transformation $^{e}\mathbf{T}_{l}$ is paramount to the process of 3D reconstruction.
-We refer to this procedure of estimation the LRF to end effector transformation, as the Lidar to end-effector extrinsic calibration.
+This procedure of estimation the LRF to PTU transformation is refered as the LRF-PTU extrinsic calibration.
 
-In this work, we propose a novel approach to perform Lidar to end-effector extrinsic calibration.
+In this work, a novel approach to perform LRF-PTU extrinsic calibration is proposed.
 The approach is based on an optimization procedure which computes the transformation that generates planar point cloud sections in areas which are marked as planes.
@@
 to the real scene geometry should be minimal.
-This paper focuses on planar geometries because of two reasons:  they are easily segmented from the full point cloud, and it is straightforward to compute a similarity metric between the point cloud and the expected planar geometry.
-In other words, in a point set representing a planar surface, the deviation from the points to the planar surface should be the lowest, if the calibration is correct.
+This method focuses on planar geometries because of two reasons: they are easily segmented from the full point cloud, and it is straightforward to compute a similarity metric between the point cloud and the expected planar geometry.
+In other words, the calibration is correct if the deviation from a point set to the corresponding planar surface is minimal.
 
-This method can then be formulated as an optimization problem: for each extrinsic calibration transformation $\mathbf{T}$, corresponds a point cloud $\mathcal{P}$.
+This method can be formulated as an optimization problem: for each extrinsic calibration transformation $\mathbf{T}$, corresponds a point cloud $\mathcal{P}$.
 This point cloud is evaluated by a cost function, which determines quantitatively how good each generated point cloud is.
@@
 
-To further explain this method, an explanation of it in each of its components will be done: the parametrization, the cost function and the optimizer.
-Each of this components are orthogonal to the others so they can be understood independently.
-
-The optimization algorithm chosen was the Powell's method, described in \cite{powell64}.
-This method finds a local minimum of a multi-dimensional unconstrained function, and does not require the gradient of this function (is unknown in this problem), which fits this particular optimization.
-This method is implemented in the python scientific library SciPy.
-However, this method requires multiple evaluations of the objective function at each iteration, in order to calculate a approximate Jacobian matrix in the neighborhood of each point.
-This usually reflect on slow optimizations, compared to other methods, such as the Gradient Descent.
+A explanation of each components will be done: the parameterization, the cost function and the optimizer. Each component is independent and can be understool separately.
 
@@
 
-The parameters in this calibration are six values that define a geometric transformation in space, which is, in the end, cast back to the transformation matrix $\bm{T}$.
+In this calibration, six parameters need to be evaluated representing the geometric transformation in space, determined by the transformation matrix $\bm{T}$.
 This transformation is decomposed into two components, a translation, and a rotation.
@@
 However, not all representations are suitable for the optimization procedure.
-In fact, in~\cite{hornegger99} the term fair parameterization was introduced: a parameterization is called fair if it does not introduce more numerical sensitivity than the one inherent to the problem itself.
-Therefore, fair parameterizations are required for optimizations, as they increase the chances for convergence.
-For example, Euler angles, which are probably the most used angle parameterizations, are not suitable~\cite{schmidt01}, because they do not yield smooth movements.
-Each rotation is non-unique and, most notably, there are singularities, so-called \textit{gimbal-lock} singularities, where one degree of freedom is lost~\cite{schmidt01}.
-Also, quaternions are not suitable for optimizations, because quaternions have $4$ components which are constrained to a unitary length.
-Despite being a fair parameterization, quaternions introduce some complexity in the algorithm to handle this constraint, so they are not usually used for optimizations~\cite{schmidt01}.
+In fact, parameterizations should not introduce more numerical sensibility than the one inherent to the problem itself.
+This parameterizations are called fair~\cite{hornegger99}.
+Common parameterizations, such as the Euler angles, and quaternions are not fair parameterizations, because of the \textit{gimbal-lock} singularities (in Euler angles), and the unitary length constrain (in quaternions)~\cite{schmidt01}.
 
-The axis/angle parameterization is the most widely used to represent a rotation in optimization procedures, as it is a fair parameterization and has only three components.
+The axis-angle representation is the most widely used to represent a rotation in optimization procedures, as it is a fair parameterization and has only three components.
 Any rotation can be represented as a rotation around an axis $a$, by an angle $\theta$.
@@
 
-This optimization was quite robust to the initial parameters, so the first guess was always a null translation and the rotation was done doing a visual inspection of the laser scanner, using angles multiples of \SI{90}{\degree}.
-
 \subsection{Cost Function}
 
-The cost function is a function used in optimization procedures that measures the difference of the result of a model and its expected result.
+The cost function is a function used in optimization procedures to measure the difference of the result of a model and its expected result.
 It then yield a value that quantitatively describes the dissimilarity between the two.
 In this case, the cost function evaluates the distance of each point $\mathbf{p}$ in the point cloud $\mathcal{P}$ to their respective plane.
-The greater this value, the greater the deformation of the point cloud, and thus the worst the calibration is.
-This objective is, then, to find the set of parameters that minimize the cost function.
+The larger this value, the larger the deformation of the point cloud, and thus the worst the calibration.
 
@@
 This segmentation is done prior to the calibration procedure, using the point cloud reconstructed using an initial estimate.
-Also, this segmentation was done manually because most segmentation algorithms, for example the RANSAC algorithm, were not capable of achieving a reliable segmentation, because the point cloud had significant deformation.
-In addition, manual segmentation is easy to do and accurate, considering that it is a one-time process.
+The initial estimate was determined by a visual inspection of the system.
 
-During the optimization, this segmentation serves as a blueprint for all segmentations.
-Each point cloud is generated in the same way, so the sequence of points is always the same.
-Therefore, it is always possible to match any point on the generated point cloud to the point in the segmented point cloud, and get the corresponding cluster index for all the points.
+The segmentation was was performed manually because it was not relevant for the scope of this work to perform it automatically.
+Also, it was a quick and a one-time process, but it was necessary to ensure that the segmentation was well performed.
+In the future, automatic segmentation procedures could be studied to streamline the method.
+
+During the optimization, the segmentation is loaded into memory as a loopup table. At each iteration, the reconstructed point cloud is segmented the same way, via this lookup table.
 
@@
 After the segmentation, the calibration procedure can be performed.
-In an initial step, the plane equation for each point cluster is computed, using the Principal Component Analysis method, or PCA.
+The plane equation for each point cluster is computed using the Principal Component Analysis method, or PCA.
 First the centroid $\bar{p}$ of each plane is found, which is the same as the mean value of all the points $p: (x, y, z) \in \textbf{R}^3$:
@@
 
-Then, the covariance matrix $\mathcal{C}$ is calculated:
+The covariance matrix $\mathcal{C}$ is then calculated:
 
@@
 
-Finally, the scores of the clusters are combined into a scalar value, which is the error of the point cloud.
-The method found was to, again, calculate the RMS of the values of the partial losses of each cluster $\sigma_i$, according to:
+The principal axes of the plane are found by the eigenvalue decomposition of the covariance matrix $\bm{C}$.
+According to the PCA method, each eigenvalue represent the statistical deviation of the points in each principal axis, more specifically, the mean square error.
+Therefore the deviation, or statistical error, of each point cluster $\bm{C}_i$ to the corresponding plane $\sigma_i$ is determined by the smallest eigenvalue $\lambda_3$.
+The cost function can the sum of all $\sigma_i$, according to:
 %
 \begin{equation}
-    \textrm{f} = \sqrt{\sum_{i}^{N}{\sigma_i^2}}.
+    \textrm{f} = \sum_{i}^{N}{\sigma_i}.
 \end{equation}
-%
-\noindent This value is expected to be minimal when all the partial losses are minimal which, according to this hypothesis, corresponds to a correct calibration.
+
+\subsection{Optimizer}
+
+The optimization algorithm chosen was the Powell's method, described in \cite{powell64}.
+This method finds a local minimum of a multi-dimensional unconstrained function, and does not require the gradient of this function (which is not known is this problem), which fits this particular optimization.
+This method is implemented in the python scientific library SciPy.
+However, this method requires multiple evaluations of the objective function at each iteration, in order to calculate a approximate Jacobian matrix in the neighborhood of each point.
+This usually results on slow optimizations, compared to other methods, such as the Gradient Descent.
 
@@
 
-In this section, the proposed method for the calibration will be evaluated using acquired data from a real scene.
-The infraestructure used to acquire the data, as well as a description of the data will be done.
-Afterward, both a qualitative and quantitative analisys of the results will be performed, to determine how much the geometry of the calibrated point cloud improved, compared to the uncalibrated point cloud.
+In this section, the proposed method for the calibration was evaluated using acquired data from a real scene.
 
 The hardware infrastructure used in this work to acquire the data was the 3D scanner \textit{lemonbot}, show in \cref{figure:lemonbot}.
-The principal components of this 3D scanner are the PTU \textit{FLIR PTU-D46} and the LRF \textit{Hokuyo UTM30LX}.
+The principal components of this 3D scanner are the PTU \textit{FLIR PTU-D46} and the 2D laser scanner \textit{Hokuyo UTM30LX}\footnote{The 3D scanner, as seen in the \cref{figure:lemonbot}, has two laser scanners: the \textit{Hokuyo UTM30LX} and the \textit{SICK LMS100}.
+In this work, only the first was used.}.
 The LRF was mounted on the PTU, to have the scan plane in the vertical direction.
-The LRF can cover \SI{270}{\degree} with an angular resolution of \SI{0.25}{\degree}.
+The laser scanner can cover \SI{270}{\degree} with an angular resolution of \SI{0.25}{\degree}.
 The PTU has an angular resolution of \SI{0.0032}{\degree}, so the error in the transformation $^{m}\mathbf{T}_{e}$ (see \cref{equation:point-reconstruction}) is negligeble.
@@
 
-For the verification, a set of 3 datasets was acquired from a single room, which corresponds to 3 point clouds.
+For the verification, a dataset was acquired from a single room, which consists of 3 acquisitions\footnote{One acquisition corresponds to a set of laser scans taken from a particular position in the scene.
+Each acquisition is processed and results in a point cloud.}.
 The acquisitions were performed in different positions in the room.
@@
 
-The optimization procedure was successful in minimizing the objective function in every dataset, as seen in \cref{table:iteration-results}.
+The optimization procedure was successful minimizing the objective function in every dataset, as seen in \cref{table:iteration-results}.
 The final results a qualitative improvement in comparison to the uncalibrated point cloud, as show in \cref{figure:visual-comparison}.
 The deformation found in the uncalibrated point cloud are almost imperceptible in the calibrated point cloud.
+
+Pratical experience shows that this calibration is resilient to the initial parameters. In this calibration, the initial parameters were roughly determined by a visual inspection of the 3D scanner.
 
@@
     \caption{Side-by-side comparison of the uncalibrated (on the left) and calibrated (on the right) point clouds.
-The most noticeable distortions are marked in red, which did not appear in the calibrated point clouds.
-This indicated that the method successfully reduced the distortions in the point clouds.}
+The most noticeable distortions are marked in the uncalibrated point clouds, and does not appear in the respective calibrated point clouds.
+This indicate that the method successfully reduced the distortions in the point clouds.}
     \label{figure:visual-comparison}
@@
 
-Despite the good qualitative result shown prior, a further quantitative analysis was performed in order to evaluate the method. The results were prepared in the following steps:
+Despite the good qualitative result, a further quantitative analysis was performed in order to evaluate the method. The results were prepared in the following steps:
 
@@
 
-This method was performed for 3 planes, and the segments can be seen in \cref{table:quantitative-results}.
+This method was performed for 3 manually segmented planes, and the segments can be seen in \cref{table:quantitative-results}.
 As seen the statistical deviation is in fact smaller in the calibrated point cloud.
@@
         \toprule
-        Point Cloud & $\mu$ (average) & $\sigma$ (std. dev.) \\
+                       & $\mu$ (average) & $\sigma$ (std. dev.) \\
         \midrule
@@
     \includegraphics[width=8cm]{images/pointclouds-histogram.png}
-    \caption{Distribution of the distances of the points to the plane. The uncalibrated points is a wider than the calibrated points, which indicates an improved calibration.}
+    \caption{Distribution of the distances of the points to the plane.
+The deviation of the uncalibrated points is a wider than the deviation of the calibrated points, which indicates an improved calibration.}
     \label{figure:deviation-histogram}
@@
 
-This paper presents a calibration procedure to determine the extrinsic transformation of LRF mounted on a robotic arm.
+This paper presents a calibration procedure to determine the extrinsic transformation of 2D-LRF and PTU.
 This transformation is uttermost necessary for multiple applications, most noticeable, 3D reconstruction systems.
@@
 
+Further work can be done in order to improve or validate further this algorithm, such as:
+
+\begin{itemize}
+    \item a study of a automatic segmentation procedure;
+    \item determation of the jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent;
+    \item validating this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm;
+    \item comparison the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}.
+\end{itemize}
+
 \bibliographystyle{IEEEtran}

### 2019-02-08 paper.tex
final revision

@@
 \usepackage{bookmark}
+\usepackage{dsfont}
 
@@
 \title{
-    2D-LRF to PTU calibration \\
-    using planar features
+    Extrinsic Calibration of 2D Laser Range Finders \\
+    using Planar Features
 }
 
-\author{\IEEEauthorblockN{Bernardo Lourenço}
-\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
-\textit{Universidade de Aveiro}\\
-Aveiro, Portugal \\
-bernardo.lourenco@ua.pt}
-\and
-\IEEEauthorblockN{Paulo Dias}
-\IEEEauthorblockA{\textit{IEETA} \\
-\textit{Universidade de Aveiro}\\
-Aveiro, Portugal \\
-paulo.dias@ua.pt}
-\and
-\IEEEauthorblockN{Miguel Riem Oliveira}
-\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
-\textit{Universidade de Aveiro}\\
-Aveiro, Portugal \\
-mriem@ua.pt}
-}
+% \author{\IEEEauthorblockN{Bernardo Lourenço}
+% \IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
+% \textit{Universidade de Aveiro}\\
+% Aveiro, Portugal \\
+% bernardo.lourenco@ua.pt}
+% \and
+% \IEEEauthorblockN{Paulo Dias}
+% \IEEEauthorblockA{\textit{IEETA} \\
+% \textit{Universidade de Aveiro}\\
+% Aveiro, Portugal \\
+% paulo.dias@ua.pt}
+% \and
+% \IEEEauthorblockN{Miguel Riem Oliveira}
+% \IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
+% \textit{Universidade de Aveiro}\\
+% Aveiro, Portugal \\
+% mriem@ua.pt}
+% }
 
@@
 \begin{abstract}
-The Laser Range Finder, or LRF, is an essential sensor in the field of robotics, providing accurate range measurements with high angular resolution.
-This sensors are also popular in 3D reconstruction systems, mounted in kinematic chains, such as PTUs.
-However, the extrinsic transformation of the LRF has to be determined with high accuracy, otherwise the 3D reconstructions will be unusable for any subsequent procedure.
-In this work, a calibration procedure which determines this transformation is described which promises high accuracy and requires no external markers.
-This method was evaluated using real 3D reconstruction data, and proved to be successful in finding the correct transformation.
+    2D Laser Range Finders, or 2D-LRFs, are essential sensors in the field of robotics, providing accurate range measurements with high angular resolution.
+These sensors can be assembled on top of systems which, by granting additional degrees of freedom to the movement of the LRF, enable the 3D reconstruction of a scene.
+The reconstruction procedure consists of the concatenation of each scan in a single point cloud representation.
+To do so, the extrinsic transformation between the LRF and the motorized system, in this case, a Pan-tilt unit, must be known with high accuracy, otherwise, the quality of the 3D reconstructed point clouds is insufficient.
+    In this work, a calibration procedure which determines this transformation is proposed.
+The method does not require a dedicated marker, which is commonly necessary and has numerous disadvantages.
+Qualitative inspections show that the proposed method is able to significantly reduce artifacts which typically appear on uncalibrated point clouds.
+Furthermore, quantitative results demonstrate that the calibrated point cloud represents the geometries present in the scene with much higher accuracy when compared with the uncalibrated point cloud.
 \end{abstract}
@@
 
-The Laser Range Finder, or LRF, is one of the most important sensors in the field of robotics, employed in applications such as Simultaneous Localization and Mapping~\cite{wang14,eyice18} and 3D reconstruction~\cite{saito10}.
-Its use is increasing due to their continuously decrease in price, new research and development and its fundamental role in emerging technologies such as autonomous driving.
+Laser Range Finder, or LRFs, is one of the most important sensors in the field of robotics, employed in applications such as Simultaneous Localization and Mapping~\cite{wang14,eyice18} and 3D reconstruction~\cite{saito10}.
+Their use is increasing due to several factors, such as decreasing costs, new research and development, and a fundamental role in emerging technologies such as autonomous driving.
 
@@
 
-This calibration is utterly necessary because any error introduces a shift on the position of each point, which in turn can be seen as deformations on the point cloud resulting from the aquisition.
-Such deformation can be seen in \cref{fig:deformed-pointcloud}, which is shown as multiple missalign planes.
-This deformations have a significant impact on subsequent point cloud processing methods, that rely on the accurate representation of the 3D environments, i.e.
+This calibration is utterly necessary because any error introduces a shift on the position of each point, which in turn can be seen as deformations on the point cloud resulting from the acquisition.
+Such deformation can be seen in \cref{fig:deformed-pointcloud}, which is shown as multiple misalign planes.
+These deformations have a significant impact on subsequent point cloud processing methods, that rely on the accurate representation of the 3D environments, i.e.
 registration (for example, ICP), segmentation and feature extraction algorithms.
@@
 In this field, there are two main approaches, depending on the sensor which the LRF is calibrated relative to.
-The first method, and the most studied, is the calibration of the LRF relative to a camera~\cite{chen16,kassir10,zhang04}, or cameras~\cite{haeselich12}.
-In this method, the static transformation between these two sensors is determined, and relies on a correspondence between visual features captured by the camera and geometric features captured by the LRF.
-This features might be markers placed on the field of view of both sensor, such as a chessboard, in the case of \cite{kassir10}.
-However, this methods have some disadvantages in the case of LRF mounted on robotic arms: first, they require the inclusion of a camera in the system; secondly, the calibration of the camera has to be performed; and lastly, the error of both calibrations is combined during the reconstruction, which increases the inaccuracies of the final reconstruction.
+The first method and the most studied is the calibration of the LRF relative to a camera~\cite{chen16,kassir10,zhang04}, or cameras~\cite{haeselich12}.
+In this method, the static transformation between these two sensors is determined and relies on a correspondence between visual features captured by the camera and geometric features captured by the LRF.
+This features might be markers placed on the field of view of both sensors, such as a chessboard, in the case of \cite{kassir10}.
+However, these methods have some disadvantages in the case of LRF mounted on robotic arms: first, they require the inclusion of a camera in the system; secondly, the calibration of the camera has to be performed; and lastly, the error of both calibrations is combined during the reconstruction, which increases the inaccuracies of the final reconstruction.
 
 Another approach is to calibrate the 2D-LRF directly against a moving platform~\cite{zeng18}.
-Most of these methods rely on the comparison between the structure of the reconstructed points with a object with a known 3D geometry, such as a plane, a ball~\cite{pereira16} or a cone~\cite{almeida12}.
+Most of these methods rely on the comparison between the structure of the reconstructed points with an object with a known 3D geometry, such as a plane, a ball~\cite{pereira16} or a cone~\cite{almeida12}.
 In~\cite{kim13}, a marker formed by two perpendicular planes is used, and their points are segmented from the reconstructed point cloud.
@@
 However, this method presents some limitations: first, ensuring that two planes are perpendicular with a small error can be very hard (it requires a well-built marker); second, it is not straight-forward how to increase the number of planes; finally, by using a small marker, only a small portion of the space is occupied, which means that only a portion of the sensors' full field of view is used, which may lead to a sub-optimal result.
-
-% TODO: colocar qualquer coisa aqui...
 
@@
 Each laser scan is a set of range measurements taken along several directions all coincident with a plane: the scanning plane.
-That is in fact why these equipment have a 2D nature.
+That is in fact why this equipment has a 2D nature.
 Thus, for each scan, a 2D slice of the scene is measured.
@@
 The process of accumulation is, in essence, based on the concatenation of 3D points from multiple scans.
-However, the grouping of these points requires that the points of each laser scan are transformed to a common, static, reference frame, the map frame.
+However, the grouping of these points requires that the points of each laser scan are transformed into a common, static, reference frame, the map frame.
 In order to achieve it, one must apply a geometric transformation to each 3D point $\mathbf{q}^{(i)}$ measured in scan $i$, represented in the LRF's local coordinate system $l$.
@@
 Thus, an accurate estimation of transformation $^{e}\mathbf{T}_{l}$ is paramount to the process of 3D reconstruction.
-This procedure of estimation the LRF to PTU transformation is refered as the LRF-PTU extrinsic calibration.
+This procedure of estimation the LRF to PTU transformation is referred to as the LRF-PTU extrinsic calibration.
 
@@
 
-A explanation of each components will be done: the parameterization, the cost function and the optimizer. Each component is independent and can be understool separately.
+An explanation of each component will be done: the parameterization, the cost function, and the optimizer. Each component is independent and can be understood separately.
 
@@
 In fact, parameterizations should not introduce more numerical sensibility than the one inherent to the problem itself.
-This parameterizations are called fair~\cite{hornegger99}.
-Common parameterizations, such as the Euler angles, and quaternions are not fair parameterizations, because of the \textit{gimbal-lock} singularities (in Euler angles), and the unitary length constrain (in quaternions)~\cite{schmidt01}.
+These parameterizations are called fair~\cite{hornegger99}.
+Common parameterizations, such as the Euler angles, and quaternions are not fair parameterizations, because of the \textit{gimbal-lock} singularities (in Euler angles), and the unitary length constraint (in quaternions)~\cite{schmidt01}.
 
@@
 The cost function is a function used in optimization procedures to measure the difference of the result of a model and its expected result.
-It then yield a value that quantitatively describes the dissimilarity between the two.
+It then yields a value that quantitatively describes the dissimilarity between the two.
 In this case, the cost function evaluates the distance of each point $\mathbf{p}$ in the point cloud $\mathcal{P}$ to their respective plane.
@@
 This segmentation is done prior to the calibration procedure, using the point cloud reconstructed using an initial estimate.
-The initial estimate was determined by a visual inspection of the system.
+The initial estimate was determined by visual inspection of the system.
 
-The segmentation was was performed manually because it was not relevant for the scope of this work to perform it automatically.
-Also, it was a quick and a one-time process, but it was necessary to ensure that the segmentation was well performed.
+The segmentation was performed manually because it was not relevant for the scope of this work to perform it automatically.
+Also, it was a quick and one-time process, but it was necessary to ensure that the segmentation was well performed.
 In the future, automatic segmentation procedures could be studied to streamline the method.
 
-During the optimization, the segmentation is loaded into memory as a loopup table. At each iteration, the reconstructed point cloud is segmented the same way, via this lookup table.
+During the optimization, the segmentation is loaded into memory as a lookup table. At each iteration, the reconstructed point cloud is segmented the same way, via this lookup table.
 
@@
 The plane equation for each point cluster is computed using the Principal Component Analysis method, or PCA.
-First the centroid $\bar{p}$ of each plane is found, which is the same as the mean value of all the points $p: (x, y, z) \in \textbf{R}^3$:
+First, the centroid $\bar{p}$ of each plane is found, which is the same as the mean value of all the points $p: (x, y, z) \in \mathds{R}^3$:
 
@@
 The covariance matrix $\mathcal{C}$ is then calculated:
-
+%
 \begin{equation}
-    \mathcal{C} = \sum_{i}{(p_i - \bar{p}) \otimes (p_i - \bar{p})} \footnote{$\otimes$ is the outer tensor product.}.
+    \mathcal{C} = \sum_{i}{(p_i - \bar{p}) \otimes (p_i - \bar{p})},
         \label{eqn:covariance-matrix}
 \end{equation}
+%
+\noindent where $\otimes$ is the outer tensor product.
 
-Then, the principal axes of the plane is find by an eigen-decomposition of the covariance matrix.
+Then, the principal axes of the plane are found by an eigendecomposition of the covariance matrix.
 The smallest eigenvalue $\lambda_3$ will be the variance $\sigma^2$ of the cluster.
@@
 So, $\sigma^2$ can be a quantitative factor to measure the cost or each cluster.
-Formally, let us admit that the $\sigma^2$ has two components: the statistical error of the laser sensor $\sigma^2_{sensor}$, which is not affected by the calibration and a second component $\sigma^2_{calib}$, which depends of the calibration error.
-Thus, the idea is that, by minimizing $\sigma^2$, a exact calibration can be obtained.
+Formally, let us admit that the $\sigma^2$ has two components: the statistical error of the laser sensor $\sigma^2_{sensor}$, which is not affected by the calibration and a second component $\sigma^2_{calib}$, which depends on the calibration error.
+Thus, the idea is that, by minimizing $\sigma^2$, an exact calibration can be obtained.
 For this calibration, however, the value $\sigma$ was used instead of $\sigma^2$, which is known as the Root Mean Square Deviation, or RMS.
@@
 The principal axes of the plane are found by the eigenvalue decomposition of the covariance matrix $\bm{C}$.
-According to the PCA method, each eigenvalue represent the statistical deviation of the points in each principal axis, more specifically, the mean square error.
+According to the PCA method, each eigenvalue represents the statistical deviation of the points in each principal axis, more specifically, the mean square error.
 Therefore the deviation, or statistical error, of each point cluster $\bm{C}_i$ to the corresponding plane $\sigma_i$ is determined by the smallest eigenvalue $\lambda_3$.
@@
 
-The optimization algorithm chosen was the Powell's method, described in \cite{powell64}.
-This method finds a local minimum of a multi-dimensional unconstrained function, and does not require the gradient of this function (which is not known is this problem), which fits this particular optimization.
+The optimization algorithm chosen was Powell's method, described in \cite{powell64}.
+This method finds a local minimum of a multi-dimensional unconstrained function and does not require the gradient of this function (which is not known is this problem), which fits this particular optimization.
 This method is implemented in the python scientific library SciPy.
-However, this method requires multiple evaluations of the objective function at each iteration, in order to calculate a approximate Jacobian matrix in the neighborhood of each point.
-This usually results on slow optimizations, compared to other methods, such as the Gradient Descent.
+However, this method requires multiple evaluations of the objective function at each iteration, in order to calculate an approximate Jacobian matrix in the neighborhood of each point.
+This usually results in slow optimizations, compared to other methods, such as Gradient Descent.
 
@@
 The laser scanner can cover \SI{270}{\degree} with an angular resolution of \SI{0.25}{\degree}.
-The PTU has an angular resolution of \SI{0.0032}{\degree}, so the error in the transformation $^{m}\mathbf{T}_{e}$ (see \cref{equation:point-reconstruction}) is negligeble.
-This 3D scanner was programmed using the \textit{ROS} framework, which is handles the hardware drivers, the transformation graph and the recording of the scanning data.
+The PTU has an angular resolution of \SI{0.0032}{\degree}, so the error in the transformation $^{m}\mathbf{T}_{e}$ (see \cref{equation:point-reconstruction}) is negligible.
+This 3D scanner was programmed using the \textit{ROS} framework, which handles the hardware drivers, the transformation graph and the recording of the scanning data.
 
@@
 The acquisitions were performed in different positions in the room.
-The room was not cluttered and the ceiling, walls and floor are all perpendicular and known to be planar surfaces.
-This was ensured in order to minimize the segmentation and calibration errors.
+The room was not cluttered and the ceiling, walls, and floor are all perpendicular and known to be planar surfaces.
+This was ensured in order to minimize segmentation and calibration errors.
 
 The optimization procedure was successful minimizing the objective function in every dataset, as seen in \cref{table:iteration-results}.
-The final results a qualitative improvement in comparison to the uncalibrated point cloud, as show in \cref{figure:visual-comparison}.
-The deformation found in the uncalibrated point cloud are almost imperceptible in the calibrated point cloud.
+The final results in a qualitative improvement in comparison to the uncalibrated point cloud, as shown in \cref{figure:visual-comparison}.
+These examples show the most common defects shown in poorly calibrated reconstructions.
+In the first example, the roof surface appears as multiple unaligned surfaces.
+This effect can be also seen in a wall, in the third example, which appears in duplication.
+The second example shows a deformation, which evidently deformed the corner structure in the point cloud.
+These deformations found in the uncalibrated point clouds are almost imperceptible in the calibrated point cloud.
 
-Pratical experience shows that this calibration is resilient to the initial parameters. In this calibration, the initial parameters were roughly determined by a visual inspection of the 3D scanner.
+Practical experience shows that this calibration is resilient to the initial parameters. In this calibration, the initial parameters were roughly determined by visual inspection of the 3D scanner.
 
@@
 
-    \caption{Side-by-side comparison of the uncalibrated (on the left) and calibrated (on the right) point clouds.
-The most noticeable distortions are marked in the uncalibrated point clouds, and does not appear in the respective calibrated point clouds.
-This indicate that the method successfully reduced the distortions in the point clouds.}
+    \caption{Side-by-side comparison of the uncalibrated (left) and calibrated (right) point clouds.
+The most noticeable distortions are marked in the uncalibrated point clouds and do not appear in the respective calibrated point clouds.
+This indicates that the method successfully reduces the distortions in the point clouds.}
     \label{figure:visual-comparison}
@@
 
-This method was performed for 3 manually segmented planes, and the segments can be seen in \cref{table:quantitative-results}.
-As seen the statistical deviation is in fact smaller in the calibrated point cloud.
-As an alternative view, in \cref{figure:deviation-histogram}, the histogram and distribution of the signed deviation is shown for both the point clouds.
+This method was performed for 3 manually segmented planes, such as the ones seen in \cref{figure:cluster-segmentation-1}, and the segments can be seen in \cref{table:quantitative-results}.
+As seen the statistical deviation is, in fact, smaller in the calibrated point cloud.
+As an alternative view, in \cref{figure:deviation-histogram}, the histogram, and distribution of the signed deviation are shown for both the point clouds.
 As shown, the deviations of the calibrated point cloud are less scattered than the deviations of the uncalibrated point cloud.
-This results indicate that this calibration method in fact improves the results.
+These results indicate that this calibration method, in fact, improves the results.
 
@@
         \toprule
-                       & $\mu$ (average) & $\sigma$ (std. dev.) \\
+                       & $\mu / \si{\meter}$ (average) & $\sigma / \si{\meter}$ (std. dev.) \\
         \midrule
@@
     \caption{Distribution of the distances of the points to the plane.
-The deviation of the uncalibrated points is a wider than the deviation of the calibrated points, which indicates an improved calibration.}
+The uncalibrated points is wider than the calibrated points, which indicates that the calibration improved the reconstructed point cloud.}
     \label{figure:deviation-histogram}
@@
 This paper presents a calibration procedure to determine the extrinsic transformation of 2D-LRF and PTU.
-This transformation is uttermost necessary for multiple applications, most noticeable, 3D reconstruction systems.
-The incorrect calibration of this systems results in artifacts and deformations on the final reconstructions, which invalidates any successive work in this point clouds.
-The proposed method is based on the measurement of the deformations of reconstructed planar surfaces, and through an optimization procedure, this deformations are minimized.
-The method was proved to be effective, through a visual inspection of the resulting point cloud and through a statistical evaluation of the reconstructed geometries.
+The accuracy of this transformation is critical for multiple applications, of which the most noticeable are 3D reconstruction systems.
+The incorrect calibration of these systems results in artifacts and deformations on the final reconstructions, which invalidates any successive work in these point clouds.
+The proposed method is based on the measurement of the deformations of reconstructed planar surfaces, and through an optimization procedure, these deformations are minimized.
+The method was proved to be effective, through a visual inspection of the resulting point cloud and through statistical evaluation of the reconstructed geometries.
 
-Further work can be done in order to improve or validate further this algorithm, such as:
-
-\begin{itemize}
-    \item a study of a automatic segmentation procedure;
-    \item determation of the jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent;
-    \item validating this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm;
-    \item comparison the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}.
-\end{itemize}
+Further work can be done in order to improve or validate further this algorithm, such as: a study of a automatic segmentation procedure; the determination of the Jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent; the validation of this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm; and the comparison of the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}.
 

### 2019-03-26 paper.tex
changes after the review

@@
 
-% \author{\IEEEauthorblockN{Bernardo Lourenço}
-% \IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
-% \textit{Universidade de Aveiro}\\
-% Aveiro, Portugal \\
-% bernardo.lourenco@ua.pt}
-% \and
-% \IEEEauthorblockN{Paulo Dias}
-% \IEEEauthorblockA{\textit{IEETA} \\
-% \textit{Universidade de Aveiro}\\
-% Aveiro, Portugal \\
-% paulo.dias@ua.pt}
-% \and
-% \IEEEauthorblockN{Miguel Riem Oliveira}
-% \IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
-% \textit{Universidade de Aveiro}\\
-% Aveiro, Portugal \\
-% mriem@ua.pt}
-% }
+\author{\IEEEauthorblockN{Bernardo Lourenço}
+\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
+\textit{Universidade de Aveiro}\\
+Aveiro, Portugal \\
+bernardo.lourenco@ua.pt}
+\and
+\IEEEauthorblockN{Paulo Dias}
+\IEEEauthorblockA{\textit{IEETA} \\
+\textit{Universidade de Aveiro}\\
+Aveiro, Portugal \\
+paulo.dias@ua.pt}
+\and
+\IEEEauthorblockN{Miguel Riem Oliveira}
+\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
+\textit{Universidade de Aveiro}\\
+Aveiro, Portugal \\
+mriem@ua.pt}
+}
 
@@
 
-Laser Range Finder, or LRFs, is one of the most important sensors in the field of robotics, employed in applications such as Simultaneous Localization and Mapping~\cite{wang14,eyice18} and 3D reconstruction~\cite{saito10}.
+Laser Range Finder, or LRFs, is one of the most important sensors in the field of robotics, employed in applications such as Simultaneous Localization and Mapping~\cite{cadena16} and 3D reconstruction~\cite{saito10}.
 Their use is increasing due to several factors, such as decreasing costs, new research and development, and a fundamental role in emerging technologies such as autonomous driving.
@@
 In this field, there are two main approaches, depending on the sensor which the LRF is calibrated relative to.
-The first method and the most studied is the calibration of the LRF relative to a camera~\cite{chen16,kassir10,zhang04}, or cameras~\cite{haeselich12}.
+The first method and the most studied is the calibration of the LRF relative to a camera~\cite{chen16,vasconcelos12}, or cameras~\cite{haeselich12}.
 In this method, the static transformation between these two sensors is determined and relies on a correspondence between visual features captured by the camera and geometric features captured by the LRF.
@@
 The 2D-LRF is a sensor that outputs several laser scans.
-Each laser scan is a set of range measurements taken along several directions all coincident with a plane: the scanning plane.
+Each laser scan is a set of range measurements taken sequentially along several directions all coincident with a plane: the scanning plane.
 That is in fact why this equipment has a 2D nature.
@@
 Through the accumulation of the several laser scans, a dense point cloud of the scene is produced.
+It is important to consider that the scan is not acquired instantaneously.
+However, if the speed of the LRF is low, its effects are negligible.
 
@@
 
-The segmentation was performed manually because it was not relevant for the scope of this work to perform it automatically.
+The segmentation was performed manually, using the software \textit{CloudCompare}\footnote{CloudCompare (\url{https://www.danielgm.net/cc/}) is a 3D point cloud and mesh processing software.}, because it was not relevant for the scope of this work to perform it automatically.
 Also, it was a quick and one-time process, but it was necessary to ensure that the segmentation was well performed.
@@
 
-Further work can be done in order to improve or validate further this algorithm, such as: a study of a automatic segmentation procedure; the determination of the Jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent; the validation of this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm; and the comparison of the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}.
+Further work can be done in order to improve or validate further this algorithm, such as: a study of a automatic segmentation procedure; the determination of the Jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent; the validation of this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm; the comparison of the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}; and study the impact of the calibration environment (for example, the number and pose of the planes and size of the room) in the calibration accuracy.
 

### 2019-03-28 paper.tex
removed hyperref to pass ieee standards

@@
 \usepackage{siunitx}
-\usepackage{hyperref}
-\usepackage[noabbrev, capitalize]{cleveref}
+\usepackage{url}
 \usepackage{bm}
@@
 \usepackage{subfig}
-\usepackage{bookmark}
 \usepackage{dsfont}
@@
 
+\IEEEoverridecommandlockouts
+\IEEEpubid{\makebox[\columnwidth]{978-1-7281-3558-8/19/\$31.00~\copyright2019 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}
+
 \maketitle
+
+\IEEEpubidadjcol
 
@@
 This calibration is utterly necessary because any error introduces a shift on the position of each point, which in turn can be seen as deformations on the point cloud resulting from the acquisition.
-Such deformation can be seen in \cref{fig:deformed-pointcloud}, which is shown as multiple misalign planes.
+Such deformation can be seen in Figure \ref{fig:deformed-pointcloud}, which is shown as multiple misalign planes.
 These deformations have a significant impact on subsequent point cloud processing methods, that rely on the accurate representation of the 3D environments, i.e.
@@
 
-The hardware infrastructure used in this work to acquire the data was the 3D scanner \textit{lemonbot}, show in \cref{figure:lemonbot}.
-The principal components of this 3D scanner are the PTU \textit{FLIR PTU-D46} and the 2D laser scanner \textit{Hokuyo UTM30LX}\footnote{The 3D scanner, as seen in the \cref{figure:lemonbot}, has two laser scanners: the \textit{Hokuyo UTM30LX} and the \textit{SICK LMS100}.
+The hardware infrastructure used in this work to acquire the data was the 3D scanner \textit{lemonbot}, shown in Figure \ref{figure:lemonbot}.
+The principal components of this 3D scanner are the PTU \textit{FLIR PTU-D46} and the 2D laser scanner \textit{Hokuyo UTM30LX}\footnote{The 3D scanner, as seen in the Figure \ref{figure:lemonbot}, has two laser scanners: the \textit{Hokuyo UTM30LX} and the \textit{SICK LMS100}.
 In this work, only the first was used.}.
@@
 The laser scanner can cover \SI{270}{\degree} with an angular resolution of \SI{0.25}{\degree}.
-The PTU has an angular resolution of \SI{0.0032}{\degree}, so the error in the transformation $^{m}\mathbf{T}_{e}$ (see \cref{equation:point-reconstruction}) is negligible.
+The PTU has an angular resolution of \SI{0.0032}{\degree}, so the error in the transformation $^{m}\mathbf{T}_{e}$ (see Equation \ref{equation:point-reconstruction}) is negligible.
 This 3D scanner was programmed using the \textit{ROS} framework, which handles the hardware drivers, the transformation graph and the recording of the scanning data.
@@
 
-The optimization procedure was successful minimizing the objective function in every dataset, as seen in \cref{table:iteration-results}.
-The final results in a qualitative improvement in comparison to the uncalibrated point cloud, as shown in \cref{figure:visual-comparison}.
+The optimization procedure was successful minimizing the objective function in every dataset, as seen in Table \ref{table:iteration-results}.
+The final results in a qualitative improvement in comparison to the uncalibrated point cloud, as shown in Figure \ref{figure:visual-comparison}.
 These examples show the most common defects shown in poorly calibrated reconstructions.
@@
 
-This method was performed for 3 manually segmented planes, such as the ones seen in \cref{figure:cluster-segmentation-1}, and the segments can be seen in \cref{table:quantitative-results}.
+This method was performed for 3 manually segmented planes, such as the ones seen in Figure \ref{figure:cluster-segmentation-1} and the segments can be seen in Table \ref{table:quantitative-results}.
 As seen the statistical deviation is, in fact, smaller in the calibrated point cloud.
-As an alternative view, in \cref{figure:deviation-histogram}, the histogram, and distribution of the signed deviation are shown for both the point clouds.
+As an alternative view, in Figure \ref{figure:deviation-histogram}, the histogram, and distribution of the signed deviation are shown for both the point clouds.
 As shown, the deviations of the calibrated point cloud are less scattered than the deviations of the uncalibrated point cloud.
@@
 
-Further work can be done in order to improve or validate further this algorithm, such as: a study of a automatic segmentation procedure; the determination of the Jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent; the validation of this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm; the comparison of the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}; and study the impact of the calibration environment (for example, the number and pose of the planes and size of the room) in the calibration accuracy.
+Future work can be done in order to improve or validate further this algorithm, such as a study of an automatic segmentation procedure; the determination of the Jacobian of the cost function, in order to reduce the iteration time of the optimization, by enabling new optimization methods, such as the gradient descent; the validation of this method with data using different 2D-LRF and different dynamic kinematic chains, such as a robotic arm; the comparison of the point cloud generated with this reconstruction method with point clouds generated with high-end commercial 3D scanners, such as the \textit{FARO Focus 3D}; and study the impact of the calibration environment (for example, the number and pose of the planes and size of the room) in the calibration accuracy.
 

### 2019-03-30 paper.tex
final adjustements, like authors filliations

@@
 
-\author{\IEEEauthorblockN{Bernardo Lourenço}
-\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
-\textit{Universidade de Aveiro}\\
-Aveiro, Portugal \\
-bernardo.lourenco@ua.pt}
-\and
-\IEEEauthorblockN{Paulo Dias}
-\IEEEauthorblockA{\textit{IEETA} \\
-\textit{Universidade de Aveiro}\\
-Aveiro, Portugal \\
-paulo.dias@ua.pt}
-\and
-\IEEEauthorblockN{Miguel Riem Oliveira}
-\IEEEauthorblockA{\textit{Departamento de Engenharia Mecânica} \\
-\textit{Universidade de Aveiro}\\
-Aveiro, Portugal \\
-mriem@ua.pt}
+\DeclareRobustCommand*{\IEEEauthorrefmark}[1]{%
+  \raisebox{0pt}[0pt][0pt]{\textsuperscript{\footnotesize #1}}%
 }
+
+\author{
+    \IEEEauthorblockN{
+        Bernardo Lourenço\IEEEauthorrefmark{1},
+        Paulo Oliveira\IEEEauthorrefmark{2,3}
+        and Miguel Oliveira\IEEEauthorrefmark{1,3,4,5}}
+    \IEEEauthorblockA{\IEEEauthorrefmark{1}Departamento de Engenharia Mecânica, Universidade de Aveiro, Portugal}
+    \IEEEauthorblockA{\IEEEauthorrefmark{2}Departamento de Electrónica, Telecomunicações e Informática, Universidade de Aveiro, Portugal}
+    \IEEEauthorblockA{\IEEEauthorrefmark{3}Institute of Electronics and Informatics Engineering of Aveiro, Universidade de Aveiro, Portugal}
+    \IEEEauthorblockA{\IEEEauthorrefmark{4}Centro de Tecnologia Mecânica e Automação (TEMA), Universidade de Aveiro, Portugal}
+    \IEEEauthorblockA{\IEEEauthorrefmark{5}Instituto de Engenharia de Sistemas e Computadores, Tecnologia e Ciência, Portugal}}
 
@@
 \maketitle
-
-\IEEEpubidadjcol
 
@@
 The 2D-LRF is a sensor that outputs several laser scans.
-Each laser scan is a set of range measurements taken sequentially along several directions all coincident with a plane: the scanning plane.
+Each laser scan is a set of range measurements taken sequentially along several directions, all coincident with a plane: the scanning plane.
 That is in fact why this equipment has a 2D nature.
@@
 Through the accumulation of the several laser scans, a dense point cloud of the scene is produced.
-It is important to consider that the scan is not acquired instantaneously.
-However, if the speed of the LRF is low, its effects are negligible.
+Note that, theoretically, the movement of the PTU may influence the measurements of the single scan.
+However, because the frequency of scanning acquisition is high, and the velocity of the joints of the PTU is low, this effect was not considered in this work.
 
